---
title: "Osa Conservation: Movement data audit"
author:
  - Chris Beirne
site: bookdown::test-bookdown
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options: 
  chunk_output_type: console
---

# Database changes

This document summarizes data contained with the movement ecology database as of  `r Sys.Date()`.

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

library(rgdal) 
library(move)
library(dplyr)
library(leaflet)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
## Load packages for google drive ----
library(googledrive)
library(purrr)
library(readxl)
library(geosphere)
library(foreach)
library(maptools)

# For when the trapping effort file is sorted
# googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
# 
# ## Find Google Drive folder 'Centre Circle Data & Info'
# data_folder <- drive_ls(path = "08 Databases + Analysis")
# data_path <- "data" 
# dir.create(data_path) 
# drive_download(as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"),
#                path="data/2_vulture_trap_deployments.xlsx")
#capture_dat <- read_excel("data/2_vulture_trap_deployments.xlsx")

# Import passcodes
MOVE_PASS <- Sys.getenv("MOVEBANK_PASSWORD")
MOVE_USE  <- Sys.getenv("MOVEBANK_USERNAME")

loginStored <- movebankLogin(username=MOVE_USE, 
                             password=MOVE_PASS)

# Get animals to check ID's in data
# Vultures
animals <-getMovebankAnimals(study=1573471517,login=loginStored)
# Ocelot
tmp <-getMovebankAnimals(study=2526574641,login=loginStored)
# Tapir
tmp2 <- getMovebankAnimals(study=1954804459,login=loginStored)
animals <- rbind(animals, tmp,tmp2)

# For some reason they are duplicated
animals[duplicated(animals)==F,]
# They vary by the field "sensor_type_id"
animals <- animals[animals$sensor_type_id==653 & is.na(animals$sensor_type_id)==F,]

# Clean up the name
animals$animalName <- paste0(sub('\\_.*', '', animals$animalName), "_", sub('\\ .*', '', animals$taxon_canonical_name))
animals$name <- sub('\\_.*', '', animals$animalName)
animals$name[animals$name=="Robin Hood"] <- "Robin.Hood"


# Sort date objects
animals$timestamp_start <- ymd_hms(animals$timestamp_start)
animals$timestamp_end <- ymd_hms(animals$timestamp_end)


#drive_auth()
## Authenticate into googledrive service account ----
## 'GOOGLE_APPLICATION_CREDENTIALS' is what we named the Github Secret that 
## contains the credential JSON file
googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
data_folder <- drive_ls(path = "08 Databases + Analysis")
data_path <- "data"
dir.create(data_path) 
```


```{r, include=F}
#Capture_database
drive_download(file = as_id("1Q8ir4HNvP2zBFwItINhc1rmBvW2MdHnJ-mUULh_Z_oc"), path = paste0("data/cap_database"), overwrite = TRUE)
cap_db <- read_excel("data/cap_database.xlsx", sheet="database")  
# Convert the times to minutes
cap_db$total_time_processing <- as.numeric(cap_db$total_time_processing)*24*60
cap_db$delay <- as.numeric(cap_db$delay)*24*60
# Replace VALUE with NA
cap_db$individual_weight[cap_db$individual_weight=="#VALUE!"] <- NA
cap_db$individual_weight <- as.numeric(cap_db$individual_weight)

# Head
cap_db$head_length<- as.numeric(cap_db$head_length)
cap_db$head_width <- as.numeric(cap_db$head_width)

# Bill
cap_db$bill_length <- as.numeric(cap_db$bill_length)
cap_db$bill_width  <- as.numeric(cap_db$bill_width)
cap_db$bill_depth  <- as.numeric(cap_db$bill_depth)

cap_db$wing_chord  <- as.numeric(cap_db$wing_chord)
cap_db$wing_flat   <- as.numeric(cap_db$wing_flat)
cap_db$wing_span   <- as.numeric(cap_db$wing_span)
cap_db$tail_length <- as.numeric(cap_db$tail_length)

cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
cap_db$body_length <- as.numeric(cap_db$body_length)
cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

# Sort bological samples
cap_db$biological_sample <- as.numeric(cap_db$biological_sample)
# Setup the categories
cap_db$species <- as.factor(cap_db$species)


# Remove empty rows
cap_db <- cap_db[is.na(cap_db$date)==F & is.na(cap_db$country)==F,]

#trap_deployments
###################################
## THIS MAY NOT BE THE RIGHT ONE ##

drive_download(file = as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"), path = paste0("data/trap_database"), overwrite = TRUE)
trap_db <- read_excel("data/trap_database.xlsx")
trap_db <- trap_db[trap_db$deployment_id!="_",]
# Add a country code
trap_db$country <- NA
trap_db$country[trap_db$latitude>0] <- "CR"
trap_db$country[trap_db$latitude<0] <- "PE"


#biological samples
drive_download(file = as_id("1kx5BYcbbDYZLnWMp1IxQySaLRQ2I4AJhTNjZfn7m4U8"), path = paste0("data/bio_database"), overwrite = TRUE)
bio_db <- read_excel("data/bio_database.xlsx", sheet="biological_samples")

# Laboratory results
drive_download(file = as_id("1GYcqjqLG2SUKMF_s4f7JMGtgOt6snzEzHtFCsiwFiiQ"), path = paste0("data/lab_database"), overwrite = TRUE)
lab_db <- read_excel("data/lab_database.xlsx", sheet="Results")


#device_database
drive_download(file = as_id("1_oa67Po8Hpmp4VH8jqlD_Sm25Fn2GPctqp7rC5EXSuo"), path = paste0("data/track_database"), overwrite = TRUE)
track_db <- read_excel("data/track_database.xlsx")

# Make one that is just the first of each unique code
track_db_1 <- track_db[!duplicated(track_db$transmitor_identifier) & is.na(track_db$deployment_start)==F,]


#pittag_database
drive_download(file = as_id("1oDWd3QYUxpSFRKkkwNLgRJmMJrxuMwqNfAM8tkN-UII"), path = paste0("data/pit_database"), overwrite = TRUE)
pit_db <- read_excel("data/pit_database.xlsx")
colnames(pit_db)[5:8] <- pit_db[1,5:8]
pit_db <- pit_db[is.na(pit_db$pit_tag_id)==F,]
# Remove blanks
pit_db <- pit_db[,1:9]

#cluster observations
drive_download(file = as_id("1KvZ2jEfv97eAJLatlw-rEic-kPitFPiUvjXcrrX7jiQ"), path = paste0("data/clus_database"), overwrite = TRUE)
clus_db <- read_excel("data/clus_database.xlsx", sheet="Data sheet")


# Get current time and last 7 days
t <- now("America/Costa_Rica")
start_t <- t-as.difftime(7,units='days')

```

### Database changes (last 7 days)

The following table shows total number of observations in each database (left) and the new records from the last 7 days (right). The green numbers show where records have been added.

```{r, include=F}
cap_db$NEW   <- difftime(cap_db$date,t)>-7
trap_db$NEW  <- difftime(trap_db$date,t)>-7
bio_db$NEW   <- difftime(bio_db$capture_date,t)>-7
track_db_1$NEW <- difftime(track_db_1$deployment_start,t)>-7
clus_db$NEW   <- difftime(clus_db$visit_date,t)>-7

list_db <- list(cap_db, trap_db, bio_db, track_db_1, clus_db)
names(list_db) <- c("vulture_capture_database",
                                    "vulture_trap_deployments",
                                    "biological_samples",
                                    "tracking_device_deployment_database",
                                    "cluster_observations_database")


# Blank dataframe to store the output
change_dat <- data.frame(name=names(list_db),
                         total_rows = NA,
                         new_rows = NA,
                         CR_total = NA,
                         CR_new = NA,
                         PE_total =NA,
                         PE_new=NA)
i <- 2
for(i in 1:length(names(list_db)))
{
   #Overall
   tmp <- list_db[[i]] %>% group_by(NEW) %>% summarize(n=n())
   change_dat$total_rows[i] <- sum(tmp$n)
   # If you have new rows
   if(nrow(tmp)>1)
   {
     change_dat$new_rows[i] <- tmp$n[2]
   } else(change_dat$new_rows[i] <- 0)
   #Country specific
   tmp <- list_db[[i]] %>% group_by(NEW, country) %>% summarize(n=n())
   # Filter any NA;s
   tmp<- tmp[is.na(tmp$country)==F,]
   # CR
   tmp_cr <- tmp[tmp$country=="CR",]
   change_dat$CR_total[i] <- sum(tmp_cr$n)
   # If you have new rows
   if(nrow(tmp_cr)>1)
   {
     change_dat$CR_new[i] <- tmp_cr$n[2]
   } else(change_dat$CR_new[i] <- 0)
   # PE
   tmp_pe <- tmp[tmp$country=="PE",]
   change_dat$PE_total[i] <- sum(tmp_pe$n)
   # If you have new rows
   if(nrow(tmp_pe)>1)
   {
     change_dat$PE_new[i] <- tmp_pe$n[2]
   } else(change_dat$PE_new[i] <- 0)

  
}


```

```{r}
table_dat <- change_dat

# Setup highlights

table_dat$new_rows = cell_spec(table_dat$new_rows, background = ifelse(table_dat$new_rows == 0, "white", "lightgreen"))
table_dat$CR_new = cell_spec(table_dat$CR_new, background = ifelse(table_dat$CR_new == 0, "white", "lightgreen"))
table_dat$PE_new = cell_spec(table_dat$PE_new, background = ifelse(table_dat$PE_new == 0, "white", "lightgreen"))


# Setup column names 
colnames(table_dat)[2:7] <- c("total", "new", "total", "new", "total", "new") 

kbl(table_dat, escape = F) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, "All locations" = 2, "Costa Rica" = 2, "Peru" = 2)) %>% 
  column_spec (c(1,3,5,7),border_left = F, border_right = T) 
```

### Database linkage check

There are several important checks to ensure our data integrity.

**Do all vulture captures have corresponding trap deployment data?**

```{r}
tmp <- cap_db[cap_db$date %in% trap_db$date,]
tmp2 <- cap_db[!(cap_db$date %in% trap_db$date),]

tmp3 <- data.frame("yes" = nrow(tmp), "no"=nrow(tmp2))

tmp3$no <- as.numeric(tmp3$no)

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)

```

The dates and locations when we captured vultures which are not in the database are as follows:

```{r}
tmp3 <- tmp2 %>% group_by(location_code, date) %>% summarize("vultures captured"=n())

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)
```


**Do all captures with biological samples have data in the `biological_samples` database?**

```{r}
tmp <- cap_db[cap_db$biological_sample>0 & is.na(cap_db$biological_sample)==F,]
tmp2 <- bio_db


# Make a key variable
tmp$link <- paste(tmp$pit_tag_id, tmp$date, sep="_")
tmp2$link <- paste(tmp2$pit_tag_id, tmp2$capture_date, sep="_")

tmp3 <- data.frame("yes" = nrow(tmp[tmp$link %in% tmp2$link,]), "no"=nrow(tmp[!(tmp$link %in% tmp2$link),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

tmp <- tmp[!(tmp$link %in% tmp2$link),]

```

The captures whose biological samples are not currently in the database are:

```{r}
kbl(tmp[,c("local_identifier", "pit_tag_id", "date")])  %>%
  kable_classic(full_width = F)
```

**Do any biological sample have duplicated codes?**

```{r}
tmp3 <- bio_db[duplicated(bio_db$sample_code)==T,]

tmp2 <- data.frame("no" = nrow(tmp2)-nrow(tmp3), "yes"=nrow(tmp3))

kbl(tmp2)  %>%
  kable_classic(full_width = F)

```

If there are any present, the samples with duplicated `sample_codes` are as follows:

```{r}
tmp2 <- bio_db[bio_db$sample_code %in% tmp3$sample_code, ]
tmp2 <- tmp2[order(tmp2$sample_code, tmp2$pit_tag_id),]

kbl(tmp2[,colnames(tmp2)!="specific_id"])  %>%
  kable_classic(full_width = F)

```

**Do all vulture captures where a device was deployed have data in `device_deployment_database`?**

```{r}
tmp <- cap_db[cap_db$transmitor_installation==1 & is.na(cap_db$transmitor_installation)==F,]
tmp2 <- track_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$transmitor_identifier %in% tmp2$transmitor_identifier,]), "no"=nrow(tmp[!(tmp$transmitor_identifier %in% tmp2$transmitor_identifier),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

```

**Do all vultures with pit-tags have an entry in the `vulture_pit_tag` database?**

```{r}
tmp <- cap_db[nchar(cap_db$pit_tag_id)>10,]
tmp2 <- pit_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$pit_tag_id %in% tmp2$pit_tag_id,]), 
                   "no"  = nrow(tmp[!(tmp$pit_tag_id %in% tmp2$pit_tag_id),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)
```

### Recaptures

To date we have had 
`r nrow(cap_db[duplicated(cap_db$pit_tag_id)==T,])` vulture recaptures. The ID's and dates are as follows:

```{r}

tmp <- cap_db[duplicated(cap_db$pit_tag_id)==T,]
tmp2 <- cap_db[cap_db$pit_tag_id %in% tmp$pit_tag_id,]
tmp2 <- tmp2[order(tmp2$pit_tag_id, tmp2$date),]
tmp <- tmp2[, c("local_identifier", "species",  "pit_tag_id", "date", "location_code")]

kbl(tmp)  %>%
  kable_classic(full_width = F)
```

Are these expected?


# Capture database

To date we have captured vultures `r nrow(cap_db)` times, deployed `r length(unique(cap_db$transmitor_identifier))` tracking devices and `r length(unique(cap_db$pit_tag_id[nchar(cap_db$pit_tag_id)>10]))` pit tags. 

The following section provides a summary of the data taken during animal captures

### Processing time

Our average processing times for the different species are as follows:

```{r}
tmp <-cap_db %>% group_by(species) %>% summarise("Mean processing time (mins)" = round(mean(total_time_processing, na.rm=T),0))

kbl(tmp)  %>%
  kable_classic(full_width = F)

```



```{r}
# Sensors deploy and sensorys not deployed

#plot(cap_db$total_time_processing, ylab="Processing time (mins)")

fig <- plot_ly(data = cap_db, x = 1:nrow(cap_db), y = ~total_time_processing, color = ~species,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(title = 'Vulture processing time', 
                      xaxis = list(title = 'Vulture capture order'), 
                      yaxis = list(title = 'Processing time (minutes)'), 
                      legend = list(title=list(text='<b> Species </b>')))




fig

```

### Morphometrics

*Weight* 

Of `r nrow(cap_db)` captures we have *`r length(cap_db$individual_weight[is.na(cap_db$individual_weight)==T])`* captures where weight was not assessed.

Of the weights recorded the distributions are as follows:

```{r}
# Weight

f1 <- plot_ly(data=cap_db, y = ~individual_weight, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(title = 'Individual weight (g)', 
                      xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Weight (g)'))

f1


```


Other:

**Head traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~head_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~head_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.2, y = 1.0,   text = "Head length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.8, y = 1,text = "Head width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

**Bill traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~bill_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~bill_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~bill_depth, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Bill length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Bill width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Bill depth (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig






```

**Wing traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~wing_chord, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~wing_flat, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~wing_span, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f4 <- plot_ly(data=cap_db, y = ~tail_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Wing chord (cm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.35, y = 1,text = "Wing flat (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.65, y = 1,text = "Wing span (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Tail length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3, f4) 
fig <- fig %>%layout(annotations = annotations) 
fig


# cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
# cap_db$body_length <- as.numeric(cap_db$body_length)
# cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

```

**Size traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~tarsus_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~body_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~MTO_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Tarsus length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Body length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Middle toe length (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

# Trap deployment database

To date we have deployed vulture traps in `r  length(unique(trap_db$location_code)) ` different locations.

The trapping locations are as follows:

```{r}
tmp <- trap_db[,c("location_code", "latitude", "longitude")]
tmp <- tmp[duplicated(tmp)==F,]


kbl(tmp)  %>%
  kable_classic(full_width = F)


```


```{r}

m <- leaflet() %>%
      # Add a satellite image layer
      addProviderTiles(providers$Stamen.TonerLite, group="Simple") %>%
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=tmp$longitude,
                         lat=tmp$latitude, 
                         popup=paste(tmp$location_code)) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Simple", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 

```

# Biological samples database

To date we have collected `r nrow(bio_db)` biological samples from `r length(unique(bio_db$pit_tag_id))` individuals. The species breakdown is as follows:

```{r}
# add a species column
bio_db$species <- NA
#cap_db$species
bio_db$species[bio_db$code_sp=="BV"] <- "coragyps_atratus"
bio_db$species[bio_db$code_sp=="KV"] <- "sarcoramphus_papa"
bio_db$species[bio_db$code_sp=="LP"] <- "leopards_pardalis"
bio_db$species[bio_db$code_sp=="TV"] <- "cathartes_aura"

tmp <- bio_db %>% group_by(species) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F)

```

Break down of the type of sample per species:

```{r}
tmp <- bio_db %>% group_by(species, tissue ) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp[, colnames(tmp)!="species"])  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp$species))

```

And the current locations of the samples:

```{r}
tmp <- bio_db %>% group_by(tissue, storage_box) %>% summarise(samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F) 

```

# Tracking device deployment database

To date we have deployed `r nrow(track_db_1)` tracking devices.

### Do all the current vital settings match?

It is important that all of our common tag types have the same settings. Is this currently the case in the database?

We first subset to the most recent deployment settings for each tag, then compare them the most important categories:

```{r}
tmp <- track_db[is.na(track_db$deployment_start)==F,]
tmp <- tmp[order(tmp$local_identifier,tmp$deployment_start, decreasing=T),]

tmp_latest <- tmp[duplicated(tmp$local_identifier)==F,]

# Convert times into something sensible
tmp_latest$`ACC On From (HH:MM)` <- round((as.numeric(tmp_latest$`ACC On From (HH:MM)`)*24),2)
tmp_latest$`ACC On Until (HH:MM)`<- round((as.numeric(tmp_latest$`ACC On Until (HH:MM)`)*24),2)

tmp_latest$`GPS On From (HH:MM)` <- substr(tmp_latest$`GPS On From (HH:MM)`,12,16)
tmp_latest$`GPS On Until (HH:MM)`<- substr(tmp_latest$`GPS On Until (HH:MM)`,12,16)

tmp_latest$`Pinger On From (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On From (HH:MM)`)*24),2)

tmp_latest$`Pinger On Until (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On Until (HH:MM)`)*24),2)


tmp2 <- tmp_latest[, c("country", "transmitor_identifier", "local_identifier", 
               "tag_model", "ACC On From (HH:MM)", "ACC On Until (HH:MM)",
               "ACC Interval (s)",
               "GPS On From (HH:MM)", "GPS On Until (HH:MM)",
               "GPS LOW RES INTERVAL (s)", "GPS HIGH RES INTERVAL (s)",
               "Pinger On From (HH:MM)", "Pinger On Until (HH:MM)",
               "Freq. (MHz)",
               "GPRS Time (HH:MM)", "GPRS Time 2 (HH:MM)"),
           ]

# Order by country and tag model
tmp2 <- tmp2[order(tmp2$country, tmp2$tag_model, tmp2$local_identifier),]

kbl(tmp2)  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp2$country))
```

# Laboratory analyses

To date we have provcess samples with the following methodologies:

```{r}

```



## Contamination - Voltammetry and atomic absorbtion (Peru)

We processed samples from 10 individuals in Peru - specifically looking for mercury and lead contamination. Assay details are scant, and we have only a single reading per individual.

The assays were performed by MC Quimicalab, and the equipment looked fairly ancient.

### Results

```{r, warning=FALSE, message=FALSE}

dat <- labs[labs$method=="voltammetry and atomic absorption" & is.na(labs$method)==F,]

dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")


fig <- plot_ly(y =dat$result,x=dat$assay_target, color=dat$assay_target, 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=dat$local_identifier) %>% 
  layout(yaxis = list(title = 'mg_kg'))

fig
```

So this analysis looks interesting - no mercury detected on feather samples, however 5 of 10 individuals having lead contamination. Levels of >1.5 mg_kg are associated with acute exposure (see Cardiel 2011 below).

The three vultures with the highest contamination levels (OC6, OC4,and OC2) are no longer returning GPS fixes. This could represent a death of the individual or the tracking device. OC1 and OC10 are still functioning (4th and 5th values). 

Splitting the lead results by species gives us:

```{r, message=F, warning=F}
fig <- plot_ly(y =dat$result[dat$assay_target=="lead"],x=dat$species[dat$assay_target=="lead"], color=dat$species[dat$assay_target=="lead"], 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=dat$local_identifier[dat$assay_target=="lead"]) %>% 
  layout(yaxis = list(title = 'mg_kg'))

fig
```

Which suggests the bulk of the contamination is in Greater-yellow headed vultures (Cathartes melambrotus) rather than king vultures (Sarcoramphus). 

**However** I have not found may other examples of this methodology in the literature. Other details of the assay were not provided, although we have some nice photos: [link](https://docs.google.com/document/d/1UezGQGasLwwy5mSU0N-JRzSgxDtR1MlY/edit?usp=sharing&ouid=101375399114641236572&rtpof=true&sd=true).

Assay detection limit unknown... so zeros perhaps do not have much value? 

**Questions** 

- How much faith can we put in this assay?

- What other information would we need from this lab to have "publishable" results?


Cardiel, Iris E., Mark A. Taggart, and Rafael Mateo. "Using Pb–Al ratios to discriminate between internal and external deposition of Pb in feathers." Ecotoxicology and Environmental Safety 74.4 (2011): 911-917.

## Contamination - Electron miscroscope (Costa Rica)

We processed samples from seven individuals from Costa Rica - all of which were king vultures. Here we have multiple replicates per individual (between three and five).

Method: The machine used was a "Scanning Electron Microscope, MEB, Brand: Hitachi, Model: S-3700N." with the samples processed by the Centro de Investigación en Estructuras Microscópicas (CIEMic) de la Universidad de Costa. 

### Results

We had a bunch of different elements analysed:

```{r}
dat <- labs[labs$method=="Scanning Electron Microscope, MEB, Brand: Hitachi, Model: S-3700N." & is.na(labs$method)==F,]

#length(unique(dat$local_identifier))

dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")


fig <- plot_ly(y =dat$result,x=dat$assay_target, color=dat$assay_target, 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=dat$local_identifier) %>% 
  layout(yaxis = list(title = 'mg_kg'))


fig
```

Lets zoom in on the individual specific results for lead and mercury. The points are mean values and the lines standard deviations:

```{r}
res <- dat %>%  group_by(local_identifier, assay_target) %>% summarize(mean=mean(result), sd=sd(result))

tmp <- res[res$assay_target=="Lead_Pb",]

fig1 <- plot_ly(data = tmp, x = ~local_identifier, y = ~mean, type = 'scatter', mode = 'markers',

        name = tmp$assay_target[1],

        error_y = ~list(array = sd,

                        color = '#000000'))%>% 
  layout(yaxis = list(title = 'Percentage composition'))


tmp <- res[res$assay_target=="Mercury_Hg",]

fig2 <- plot_ly(data = tmp, x = ~local_identifier, y = ~mean, type = 'scatter', mode = 'markers',

        name = tmp$assay_target[1],

        error_y = ~list(array = sd,

                        color = '#000000'))%>% 
  layout(yaxis = list(title = 'Percentage composition'))


subplot(fig1, fig2, shareY=T)


```

Attenborough is still alive and well! Interestingly, the patterns look the same between both elements. The value of repeated measures through time here would be clear.

**Questions** 

- Is it possible to calculate the magnitude of contamination using molecular weights? (convert from percentages)

- Are any of the other elements of interest?

- Is it an issue that we are not using the (more complex) total and methyl-mercury assays?


## Blood cell counts

We had blood cell parameters quantified for 5 individuals in Costa Rica. 

### Results

Differentials first:

```{r}
dat <- labs[labs$method=="Cell Count" & is.na(labs$method)==F & labs$units=="%",]
table(dat$units)
#length(unique(dat$local_identifier))

dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
params <- dat$assay_target

fig <- plot_ly(y =dat$result,x=dat$assay_target, color=dat$assay_target, 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=dat$local_identifier) %>% 
  layout(yaxis = list(title = 'mg_kg'))


fig
```

Attenborough has a thrombocyte percentage outside of the range of the other data - this vulture also had high Pb and Hg levels. 

Others:

```{r}
dat <- labs[labs$method=="Cell Count" & is.na(labs$method)==F & labs$units!="%" & !(labs$assay_target %in% params),]

dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
tmp <- dat[dat$assay_target=="Erythrocytes",]
fig1 <- plot_ly(y =tmp$result,x=tmp$assay_target,
                name=paste(tmp$assay_target[1], tmp$units[1]),
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))

tmp <- dat[dat$assay_target=="Hemoglobin",]
fig2 <- plot_ly(y =tmp$result,x=tmp$assay_target,
                name=paste(tmp$assay_target[1], tmp$units[1]), 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))

tmp <- dat[dat$assay_target=="VCM",]
fig3 <- plot_ly(y =tmp$result,x=tmp$assay_target,
                name=paste(tmp$assay_target[1], tmp$units[1]), 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))

tmp <- dat[dat$assay_target=="MCC",]
fig4 <- plot_ly(y =tmp$result,x=tmp$assay_target, 
                name=paste(tmp$assay_target[1], tmp$units[1]),
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))

tmp <- dat[dat$assay_target=="HMC",]
fig5 <- plot_ly(y =tmp$result,x=tmp$assay_target, 
                name=paste(tmp$assay_target[1], tmp$units[1]),
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))


tmp <- dat[dat$assay_target=="leucocytes",]
fig6 <- plot_ly(y =tmp$result,x=tmp$assay_target, 
                name=paste(tmp$assay_target[1], tmp$units[1]),
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))

tmp <- dat[dat$assay_target=="thrombocytes",]
fig7 <- plot_ly(y =tmp$result,x=tmp$assay_target, 
                name=paste(tmp$assay_target[1], tmp$units[1]), 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))
tmp <- dat[dat$assay_target=="protein_total",]
fig8 <- plot_ly(y =tmp$result,x=tmp$assay_target,
                name=paste(tmp$assay_target[1], tmp$units[1]), 
               type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8, hovertext=tmp$local_identifier) %>% 
  layout(yaxis = list(title = tmp$units[1], rangemode="tozero"))


subplot(fig1,fig2,fig3,fig4,
        fig5,fig6,fig7,fig8, nrows=2)

```

Interestingly, Attenborough again also has elevated leucocytes and thrombocytes. The others all seems to be in a reasonable range. Could this be a cost of an exposure event? 


**Question** Are there any specific categories we should keep an eye on?

**Question** Is there a database for other bird species?













