---
title: "Osa Conservation: Movement data audit"
author:
  - Chris Beirne
site: bookdown::test-bookdown
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options: 
  chunk_output_type: console
---

# Database changes

This document summarizes data contained with the movement ecology database as of  `r Sys.Date()`.

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

library(rgdal) 
library(move)
library(dplyr)
library(leaflet)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
## Load packages for google drive ----
library(googledrive)
library(purrr)
library(readxl)
library(geosphere)
library(foreach)
library(maptools)

# For when the trapping effort file is sorted
# googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
# 
# ## Find Google Drive folder 'Centre Circle Data & Info'
# data_folder <- drive_ls(path = "08 Databases + Analysis")
# data_path <- "data" 
# dir.create(data_path) 
# drive_download(as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"),
#                path="data/2_vulture_trap_deployments.xlsx")
#capture_dat <- read_excel("data/2_vulture_trap_deployments.xlsx")

# Import passcodes
MOVE_PASS <- Sys.getenv("MOVEBANK_PASSWORD")
MOVE_USE  <- Sys.getenv("MOVEBANK_USERNAME")

loginStored <- movebankLogin(username=MOVE_USE, 
                             password=MOVE_PASS)

# Get animals to check ID's in data
# Vultures
animals <-getMovebankAnimals(study=1573471517,login=loginStored)
# Ocelot
tmp <-getMovebankAnimals(study=2526574641,login=loginStored)
# Tapir
tmp2 <- getMovebankAnimals(study=1954804459,login=loginStored)
animals <- rbind(animals, tmp,tmp2)

# For some reason they are duplicated
animals[duplicated(animals)==F,]
# They vary by the field "sensor_type_id"
animals <- animals[animals$sensor_type_id==653 & is.na(animals$sensor_type_id)==F,]

# Clean up the name
animals$animalName <- paste0(sub('\\_.*', '', animals$animalName), "_", sub('\\ .*', '', animals$taxon_canonical_name))
animals$name <- sub('\\_.*', '', animals$animalName)
animals$name[animals$name=="Robin Hood"] <- "Robin.Hood"


# Sort date objects
animals$timestamp_start <- ymd_hms(animals$timestamp_start)
animals$timestamp_end <- ymd_hms(animals$timestamp_end)


#drive_auth()
## Authenticate into googledrive service account ----
## 'GOOGLE_APPLICATION_CREDENTIALS' is what we named the Github Secret that 
## contains the credential JSON file
googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
data_folder <- drive_ls(path = "08 Databases + Analysis")
data_path <- "data"
dir.create(data_path) 
```


```{r, include=F}
#Capture_database
drive_download(file = as_id("1Q8ir4HNvP2zBFwItINhc1rmBvW2MdHnJ-mUULh_Z_oc"), path = paste0("data/cap_database"), overwrite = TRUE)
cap_db <- read_excel("data/cap_database.xlsx", sheet="database")  
# Convert the times to minutes
cap_db$total_time_processing <- as.numeric(cap_db$total_time_processing)*24*60
cap_db$delay <- as.numeric(cap_db$delay)*24*60
# Replace VALUE with NA
cap_db$individual_weight[cap_db$individual_weight=="#VALUE!"] <- NA
cap_db$individual_weight <- as.numeric(cap_db$individual_weight)

# Head
cap_db$head_length<- as.numeric(cap_db$head_length)
cap_db$head_width <- as.numeric(cap_db$head_width)

# Bill
cap_db$bill_length <- as.numeric(cap_db$bill_length)
cap_db$bill_width  <- as.numeric(cap_db$bill_width)
cap_db$bill_depth  <- as.numeric(cap_db$bill_depth)

cap_db$wing_chord  <- as.numeric(cap_db$wing_chord)
cap_db$wing_flat   <- as.numeric(cap_db$wing_flat)
cap_db$wing_span   <- as.numeric(cap_db$wing_span)
cap_db$tail_length <- as.numeric(cap_db$tail_length)

cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
cap_db$body_length <- as.numeric(cap_db$body_length)
cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

# Sort bological samples
cap_db$biological_sample <- as.numeric(cap_db$biological_sample)
# Setup the categories
cap_db$species <- as.factor(cap_db$species)


# Remove empty rows
cap_db <- cap_db[is.na(cap_db$date)==F & is.na(cap_db$country)==F,]

unique_db <- cap_db[,c("pit_tag_id", "local_identifier", "code_sp")]
# Remove NA's
unique_db <- unique_db[unique_db$local_identifier!="NA",]
# Remove duplicates
unique_db <- unique_db[duplicated(unique_db)==F,]
table(unique_db$local_identifier)

#trap_deployments
###################################
## THIS MAY NOT BE THE RIGHT ONE ##

drive_download(file = as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"), path = paste0("data/trap_database"), overwrite = TRUE)
trap_db <- read_excel("data/trap_database.xlsx")
trap_db <- trap_db[trap_db$deployment_id!="_",]
# Add a country code
trap_db$country <- NA
trap_db$country[trap_db$latitude>0] <- "CR"
trap_db$country[trap_db$latitude<0] <- "PE"


#biological samples
drive_download(file = as_id("1kx5BYcbbDYZLnWMp1IxQySaLRQ2I4AJhTNjZfn7m4U8"), path = paste0("data/bio_database"), overwrite = TRUE)
bio_db <- read_excel("data/bio_database.xlsx", sheet="biological_samples")
# Remove NA's
bio_db <- bio_db[is.na(bio_db$capture_date)==F,]

# Laboratory results
drive_download(file = as_id("1GYcqjqLG2SUKMF_s4f7JMGtgOt6snzEzHtFCsiwFiiQ"), path = paste0("data/lab_database"), overwrite = TRUE)
lab_db <- read_excel("data/lab_database.xlsx", sheet="Results")


# Laboratory results
drive_download(file = as_id("1GYcqjqLG2SUKMF_s4f7JMGtgOt6snzEzHtFCsiwFiiQ"), path = paste0("data/lab_database"), overwrite = TRUE)
lab_db <- read_excel("data/lab_database.xlsx", sheet="Results")


#device_database
drive_download(file = as_id("1_oa67Po8Hpmp4VH8jqlD_Sm25Fn2GPctqp7rC5EXSuo"), path = paste0("data/track_database"), overwrite = TRUE)
track_db <- read_excel("data/track_database.xlsx")

# Make one that is just the first of each unique code
track_db_1 <- track_db[!duplicated(track_db$transmitor_identifier) & is.na(track_db$deployment_start)==F,]


#pittag_database
drive_download(file = as_id("1oDWd3QYUxpSFRKkkwNLgRJmMJrxuMwqNfAM8tkN-UII"), path = paste0("data/pit_database"), overwrite = TRUE)
pit_db <- read_excel("data/pit_database.xlsx")
colnames(pit_db)[5:8] <- pit_db[1,5:8]
pit_db <- pit_db[is.na(pit_db$pit_tag_id)==F,]
# Remove blanks
pit_db <- pit_db[,1:9]

#cluster observations
drive_download(file = as_id("1KvZ2jEfv97eAJLatlw-rEic-kPitFPiUvjXcrrX7jiQ"), path = paste0("data/clus_database"), overwrite = TRUE)
clus_db <- read_excel("data/clus_database.xlsx", sheet="Data sheet")


# Get current time and last 7 days
t <- now("America/Costa_Rica")
start_t <- t-as.difftime(7,units='days')

```

### Database changes (last 7 days)

The following table shows total number of observations in each database (left) and the new records from the last 7 days (right). The green numbers show where records have been added.

*Note* the changes are from when the data were collected, not when they were entered in Google Drive. If they were entered a long time after the collection date they may not be shown here. 

```{r, include=F}
cap_db$NEW   <- difftime(cap_db$date,t)>-7
trap_db$NEW  <- difftime(trap_db$date,t)>-7
bio_db$NEW   <- difftime(bio_db$capture_date,t)>-7
track_db_1$NEW <- difftime(track_db_1$deployment_start,t)>-7
clus_db$NEW   <- difftime(clus_db$visit_date,t)>-7

list_db <- list(cap_db, trap_db, bio_db, track_db_1, clus_db)
names(list_db) <- c("vulture_capture_database",
                                    "vulture_trap_deployments",
                                    "biological_samples",
                                    "tracking_device_deployment_database",
                                    "cluster_observations_database")


# Blank dataframe to store the output
change_dat <- data.frame(name=names(list_db),
                         total_rows = NA,
                         new_rows = NA,
                         CR_total = NA,
                         CR_new = NA,
                         PE_total =NA,
                         PE_new=NA)
i <- 2
for(i in 1:length(names(list_db)))
{
   #Overall
   tmp <- list_db[[i]] %>% group_by(NEW) %>% summarize(n=n())
   change_dat$total_rows[i] <- sum(tmp$n)
   # If you have new rows
   if(nrow(tmp)>1)
   {
     change_dat$new_rows[i] <- tmp$n[2]
   } else(change_dat$new_rows[i] <- 0)
   #Country specific
   tmp <- list_db[[i]] %>% group_by(NEW, country) %>% summarize(n=n())
   # Filter any NA;s
   tmp<- tmp[is.na(tmp$country)==F,]
   # CR
   tmp_cr <- tmp[tmp$country=="CR",]
   change_dat$CR_total[i] <- sum(tmp_cr$n)
   # If you have new rows
   if(nrow(tmp_cr)>1)
   {
     change_dat$CR_new[i] <- tmp_cr$n[2]
   } else(change_dat$CR_new[i] <- 0)
   # PE
   tmp_pe <- tmp[tmp$country=="PE",]
   change_dat$PE_total[i] <- sum(tmp_pe$n)
   # If you have new rows
   if(nrow(tmp_pe)>1)
   {
     change_dat$PE_new[i] <- tmp_pe$n[2]
   } else(change_dat$PE_new[i] <- 0)

  
}


```

```{r}
table_dat <- change_dat

# Setup highlights

table_dat$new_rows = cell_spec(table_dat$new_rows, background = ifelse(table_dat$new_rows == 0, "white", "lightgreen"))
table_dat$CR_new = cell_spec(table_dat$CR_new, background = ifelse(table_dat$CR_new == 0, "white", "lightgreen"))
table_dat$PE_new = cell_spec(table_dat$PE_new, background = ifelse(table_dat$PE_new == 0, "white", "lightgreen"))


# Setup column names 
colnames(table_dat)[2:7] <- c("total", "new", "total", "new", "total", "new") 

kbl(table_dat, escape = F) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, "All locations" = 2, "Costa Rica" = 2, "Peru" = 2)) %>% 
  column_spec (c(1,3,5,7),border_left = F, border_right = T) 
```

### Database linkage checks

There are several important checks to ensure our data integrity - each data entry sheet links to others (e.g. capture data to biological samples).

**Do all vulture captures have corresponding trap deployment data?**

```{r}
tmp <- cap_db[cap_db$date %in% trap_db$date,]
tmp2 <- cap_db[!(cap_db$date %in% trap_db$date),]

tmp3 <- data.frame("yes" = nrow(tmp), "no"=nrow(tmp2))

tmp3$no <- as.numeric(tmp3$no)

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)

```

The dates and locations when we captured vultures which are not in the database are as follows:

```{r}
tmp3 <- tmp2 %>% group_by(location_code, date) %>% summarize("vultures captured"=n())

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)
```

&nbsp;

**Do all captures with biological samples have data in the `biological_samples` database?**

```{r}
tmp <- cap_db[cap_db$biological_sample>0 & is.na(cap_db$biological_sample)==F,]
tmp2 <- bio_db


# Make a key variable
tmp$link <- paste(tmp$pit_tag_id, tmp$date, sep="_")
tmp2$link <- paste(tmp2$pit_tag_id, tmp2$capture_date, sep="_")

tmp3 <- data.frame("yes" = nrow(tmp[tmp$link %in% tmp2$link,]), "no"=nrow(tmp[!(tmp$link %in% tmp2$link),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

tmp <- tmp[!(tmp$link %in% tmp2$link),]

```

The captures whose biological samples are not currently in the database are:

```{r}
kbl(tmp[,c("local_identifier", "pit_tag_id", "date")])  %>%
  kable_classic(full_width = F)
```
&nbsp;

**Do any biological samples have duplicated codes?**

```{r}
tmp3 <- bio_db[duplicated(bio_db$sample_code)==T,]

tmp2 <- data.frame("no" = nrow(tmp2)-nrow(tmp3), "yes"=nrow(tmp3))

kbl(tmp2)  %>%
  kable_classic(full_width = F)

tmp2 <- bio_db[bio_db$sample_code %in% tmp3$sample_code, ]
tmp2 <- tmp2[order(tmp2$sample_code, tmp2$pit_tag_id),]

```

If there are any present, the samples with duplicated `sample_codes` are shown below:

```{r}
if(nrow(tmp2)>0){
kbl(tmp2[,colnames(tmp2)!="specific_id"])  %>%
  kable_classic(full_width = F)
}else{print("No samples duplicated")}
```
&nbsp;

**Do all vulture captures where a device was deployed have data in `device_deployment_database`?**

```{r}
tmp <- cap_db[cap_db$transmitor_installation==1 & is.na(cap_db$transmitor_installation)==F,]
tmp2 <- track_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$transmitor_identifier %in% tmp2$transmitor_identifier,]), "no"=nrow(tmp[!(tmp$transmitor_identifier %in% tmp2$transmitor_identifier),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

```
&nbsp;

**Do all vultures with pit-tags have an entry in the `vulture_pit_tag` database?**

```{r}
tmp <- cap_db[nchar(cap_db$pit_tag_id)>10,]
tmp2 <- pit_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$pit_tag_id %in% tmp2$pit_tag_id,]), 
                   "no"  = nrow(tmp[!(tmp$pit_tag_id %in% tmp2$pit_tag_id),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)
```
&nbsp;

**Are all vultures in the `device_deployment_database` represented in Movebank?**

And if any are not, what are their names?

```{r}
tmp <- unique(track_db$local_identifier) 
tmp <- tmp[is.na(tmp)==F] 
tmp2 <- table(tmp %in% animals$local_identifier)

kbl(tmp2)  %>%
  kable_classic(full_width = F)

print(tmp[!(tmp %in% animals$local_identifier)])

```
&nbsp;

### Recaptures

To date we have had 
`r nrow(cap_db[duplicated(cap_db$pit_tag_id)==T,])` vulture recaptures. The ID's and dates are as follows:

```{r}

tmp <- cap_db[duplicated(cap_db$pit_tag_id)==T,]
tmp2 <- cap_db[cap_db$pit_tag_id %in% tmp$pit_tag_id,]
tmp2 <- tmp2[order(tmp2$pit_tag_id, tmp2$date),]
tmp <- tmp2[, c("local_identifier", "species",  "pit_tag_id", "date", "location_code")]

kbl(tmp)  %>%
  kable_classic(full_width = F)
```

Are these expected?


# Capture database

To date we have captured vultures `r nrow(cap_db)` times, deployed `r length(unique(cap_db$transmitor_identifier))` tracking devices and `r length(unique(cap_db$pit_tag_id[nchar(cap_db$pit_tag_id)>10]))` pit tags. 

These captures occurred on the following dates:

```{r}

tmp <- cap_db
tmp$country <- as.factor(tmp$country)
tmp$location <- as.numeric(tmp$country)
tmp$local_identifier[tmp$local_identifier=="NA"] <- tmp$pit_tag_id[tmp$local_identifier=="NA"] 
plot_ly(data = tmp, x = tmp$date, y=tmp$location, color = ~species,
               hoverinfo = 'text',height=200,
               text = ~paste(date,local_identifier,sep = "<br>")) %>% 
              layout(title = '', 
                      xaxis = list(title = 'Vulture capture date'), 
                      yaxis = list(title = "",
                                   ticktext=c("Costa Rica", "Peru"),
                                   tickvals=list(1,2)), 
                      legend = list(title=list(text='<b> Species </b>')))

```


The following section provides a summary of the data taken during animal captures

### Processing time

Our average processing times for the different species are as follows:

```{r}
tmp <-cap_db %>% group_by(species) %>% summarise("Mean processing time (mins)" = round(mean(total_time_processing, na.rm=T),0))

kbl(tmp)  %>%
  kable_classic(full_width = F)

```



```{r}
# Sensors deploy and sensorys not deployed

#plot(cap_db$total_time_processing, ylab="Processing time (mins)")

fig <- plot_ly(data = cap_db, x = 1:nrow(cap_db), y = ~total_time_processing, color = ~species,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(title = 'Vulture processing time', 
                      xaxis = list(title = 'Vulture capture order'), 
                      yaxis = list(title = 'Processing time (minutes)'), 
                      legend = list(title=list(text='<b> Species </b>')))




fig

```

### Crop size

```{r}
cap_db$crop <- as.numeric(cap_db$crop)
tmp <- as.data.frame.matrix(table(cap_db$species,cap_db$crop))

kbl(tmp) %>%
  kable_classic(full_width = F) %>%   add_header_above(header = c("Species" = 1, "crop" = 3))

```

### Morphometrics

*Weight* 

Of `r nrow(cap_db)` captures we have *`r length(cap_db$individual_weight[is.na(cap_db$individual_weight)==T])`* captures where weight was not assessed.

Of the weights recorded the distributions are as follows:

```{r}
# Weight

f1 <- plot_ly(data=cap_db, y = ~individual_weight, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(title = 'Individual weight (g)', 
                      xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Weight (g)'))

f1


```


Other:

**Head traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~head_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~head_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.2, y = 1.0,   text = "Head length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.8, y = 1,text = "Head width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

**Bill traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~bill_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~bill_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~bill_depth, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Bill length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Bill width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Bill depth (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig






```

**Wing traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~wing_chord, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~wing_flat, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~wing_span, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f4 <- plot_ly(data=cap_db, y = ~tail_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Wing chord (cm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.35, y = 1,text = "Wing flat (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.65, y = 1,text = "Wing span (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Tail length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3, f4) 
fig <- fig %>%layout(annotations = annotations) 
fig


# cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
# cap_db$body_length <- as.numeric(cap_db$body_length)
# cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

```

**Size traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~tarsus_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~body_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~MTO_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Tarsus length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Body length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Middle toe length (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

# Trap deployment database

To date we have deployed vulture traps in `r  length(unique(trap_db$location_code)) ` different locations.

The trapping locations are as follows:

```{r}
tmp <- trap_db[,c("location_code", "latitude", "longitude")]
tmp <- tmp[duplicated(tmp)==F,]


kbl(tmp)  %>%
  kable_classic(full_width = F)


```


```{r}

m <- leaflet() %>%
      fitBounds(-70,-12.8,-84,8.8) %>% 
      # Add a satellite image layer
      addProviderTiles(providers$Stamen.TonerLite, group="Simple") %>%
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=tmp$longitude,
                         lat=tmp$latitude, 
                         popup=paste(tmp$location_code)) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Simple", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 

```

# Biological samples database

To date we have collected `r nrow(bio_db)` biological samples from `r length(unique(bio_db$pit_tag_id))` individuals. The species breakdown is as follows:

```{r}
# add a species column
bio_db$species <- NA
#cap_db$species
bio_db$species[bio_db$code_sp=="BV"] <- "coragyps_atratus"
bio_db$species[bio_db$code_sp=="KV"] <- "sarcoramphus_papa"
bio_db$species[bio_db$code_sp=="LP"] <- "leopards_pardalis"
bio_db$species[bio_db$code_sp=="TV"] <- "cathartes_aura"

tmp <- bio_db %>% group_by(species) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F)

```

Break down of the type of sample per species:

```{r}
tmp <- bio_db %>% group_by(species, tissue ) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp[, colnames(tmp)!="species"])  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp$species))

```

And the current locations of the samples:

```{r}
tmp <- bio_db %>% group_by(tissue, storage_box) %>% summarise(samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F) 

```

# Tracking device deployment database

To date we have deployed `r nrow(track_db_1)` tracking devices.

### Do all the current vital settings match?

It is important that all of our common tag types have the same settings. Is this currently the case in the database?

We first subset to the most recent deployment settings for each tag, then compare them the most important categories:

```{r}
tmp <- track_db[is.na(track_db$deployment_start)==F,]
tmp <- tmp[order(tmp$local_identifier,tmp$deployment_start, decreasing=T),]

tmp_latest <- tmp[duplicated(tmp$local_identifier)==F,]

# Convert times into something sensible
tmp_latest$`ACC On From (HH:MM)` <- round((as.numeric(tmp_latest$`ACC On From (HH:MM)`)*24),2)
tmp_latest$`ACC On Until (HH:MM)`<- round((as.numeric(tmp_latest$`ACC On Until (HH:MM)`)*24),2)

tmp_latest$`GPS On From (HH:MM)` <- substr(tmp_latest$`GPS On From (HH:MM)`,12,16)
tmp_latest$`GPS On Until (HH:MM)`<- substr(tmp_latest$`GPS On Until (HH:MM)`,12,16)

tmp_latest$`Pinger On From (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On From (HH:MM)`)*24),2)

tmp_latest$`Pinger On Until (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On Until (HH:MM)`)*24),2)


tmp2 <- tmp_latest[, c("country", "transmitor_identifier", "local_identifier", 
               "tag_model", "ACC On From (HH:MM)", "ACC On Until (HH:MM)",
               "ACC Interval (s)",
               "GPS On From (HH:MM)", "GPS On Until (HH:MM)",
               "GPS LOW RES INTERVAL (s)", "GPS HIGH RES INTERVAL (s)",
               "Pinger On From (HH:MM)", "Pinger On Until (HH:MM)",
               "Freq. (MHz)",
               "GPRS Time (HH:MM)", "GPRS Time 2 (HH:MM)"),
           ]

# Order by country and tag model
tmp2 <- tmp2[order(tmp2$country, tmp2$tag_model, tmp2$local_identifier),]

kbl(tmp2)  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp2$country))%>%
    scroll_box(width = "100%", height=600px)


```

# Laboratory analyses

We first need to check if all biological samples have associated pit-tag codes, and then add in the `local_identifier` information, if it is available.

**Do all biological samples have associated pit-tag codes?**

```{r}
# Add in the pittag code
lab_full <- left_join(lab_db, bio_db[,c("sample_code", "pit_tag_id", "code_sp")])
# And if individuals have a unique ID - add it
lab_full <- left_join(lab_full, unique_db)

kbl(table(is.na(lab_full$pit_tag_id)==F)) %>%
  kable_classic(full_width = F)

```

To date we have processed samples with the following methodologies:

```{r}

tmp <- lab_full %>% group_by(assay) %>% summarise(total_assays=n(), individuals=n_distinct(pit_tag_id))

kbl(tmp) %>%
  kable_classic(full_width = F)

```

### Sex

So far we have sexed `r nrow(lab_full[lab_full$assay=="sex",])` vultures:

```{r}
tmp <- lab_full[lab_full$assay=="sex",]
tmp <- tmp[,c("local_identifier", "result", "code_sp")]
tmp <- tmp[order(tmp$code_sp,tmp$result, tmp$local_identifier),]
kbl(tmp) %>%
  kable_classic(full_width = F) 
```

### Blood cell counts

```{r}
tmp <- lab_full[lab_full$assay=="blood",]
```

We have blood cell parameters quantified for `r length(unique(tmp$pit_tag_id))` individuals and `r length(unique(tmp$code_sp))` species. 

```{r}
tmp2 <- tmp %>% group_by(code_sp) %>% summarise(individuals=n_distinct(pit_tag_id))

kbl(tmp2) %>%
  kable_classic(full_width = F) 
```

Differentials first:

```{r}
#table(tmp$units)
tmp$result <- as.numeric(tmp$result)


tmp2 <- tmp[tmp$assay_target %in% c("basophils", "eosinophils", "Hematocrit", "heterophils", "lymphocytes", "monocytes"),]
tmp2 <- tmp2[tmp2$units=="%",]
#length(unique(dat$local_identifier))
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
# Make the local identifier name blank is NA
tmp2$local_identifier[is.na(tmp2$local_identifier)==T] <- tmp2$pit_tag_id[is.na(tmp2$local_identifier)==T]

params <- tmp2$assay_target

fig <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp, 
               type = "box", boxpoints = "all", jitter = 0.1,
        pointpos = 0, hovertext=paste(tmp2$local_identifier)) %>% 
  layout(yaxis = list(title = '"% composition"'), boxmode="group")

fig
```

And the other parameters:

```{r}
dat <- tmp[tmp$method=="Cell Count" & is.na(tmp$method)==F & tmp$units!="%" & !(tmp$assay_target %in% params),]
dat$local_identifier[is.na(dat$local_identifier)==T] <- dat$pit_tag_id[is.na(dat$local_identifier)==T]
dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
tmp2 <- dat[dat$assay_target%in% c("Erythrocytes","proportion_H:L", "thrombocytes", "leucocytes", "Hemoglobin") ,]

fig1 <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp,
                 type = "box", boxpoints = "all", jitter = 0.2,
        pointpos = 0, hovertext=tmp2$local_identifier) %>% 
  layout(yaxis = list(title = "units", rangemode="tozero"), boxmode="group")

fig1

tmp2 <- dat[!(dat$assay_target%in% c("Erythrocytes","proportion_H:L", "thrombocytes", "leucocytes", "Hemoglobin")) ,]

fig1 <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp,
                 type = "box", boxpoints = "all", jitter = 0.2,
        pointpos = 0, hovertext=tmp2$local_identifier) %>% 
  layout(yaxis = list(title = "units", rangemode="tozero"), boxmode="group")

fig1
```

# Cluster visits

To date we have visited `r nrow(clus_db)' cluster sites.

The visits dates are as follows, each point represents a visit:

```{r}
clus_db$country <- as.factor(clus_db$country)
clus_db$location <- as.numeric(clus_db$country)

plot_ly(data = clus_db, x = clus_db$visit_date, y=clus_db$location,
               hoverinfo = 'text',
               text = ~paste(visit_date,bird_id,sep = "<br>"), height=200) %>% 
              layout(title = 'Cluster visits', 
                      xaxis = list(title = 'Dates'), 
                      yaxis = list(title = "",
                                   ticktext=c("Costa Rica", "Peru"),
                                   tickvals=list(1,2)))



```

Where are these clusters?

```{r}
m <- leaflet() %>%
      # Add a satellite image layer
      addProviderTiles(providers$Stamen.TonerLite, group="Simple") %>%
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=clus_db$cluster_longitude,
                         lat=clus_db$cluster_latitude, 
                         popup=paste(clus_db$visit_date)) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Simple", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 
```

Of the clusters visited so far, the type of carrion found was as follows (note NA's = no carcass).

```{r}
kbl(table(clus_db$carcass_id))%>%
  kable_classic(full_width = F) 
```

On how many occasions have we observed a bird with a GPS device feeding?

```{r}
kbl(table(clus_db$focal_bird_observed))%>%
  kable_classic(full_width = F) 

```

Who were the birds observed?

```{r}
kbl(table(clus_db$bird_id[clus_db$focal_bird_observed=="yes"]))%>%
  kable_classic(full_width = F) 
```
