---
title: "Osa Conservation: Movement data audit"
author:
  - Chris Beirne
site: bookdown::test-bookdown
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options: 
  chunk_output_type: console
---

# Database changes

This document summarizes data contained with the movement ecology database as of  `r Sys.Date()`.

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

#library(rgdal) 
library(move)
library(dplyr)
library(leaflet)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
## Load packages for google drive ----
library(googledrive)
library(purrr)
library(readxl)
library(geosphere)
library(foreach)
#library(maptools)
library(MetBrewer)
library(htmltools)
library(tidyr)

# For when the trapping effort file is sorted
# googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
# 
# ## Find Google Drive folder 'Centre Circle Data & Info'
# data_folder <- drive_ls(path = "08 Databases + Analysis")
# data_path <- "data" 
# dir.create(data_path) 
# drive_download(as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"),
#                path="data/2_vulture_trap_deployments.xlsx")
#capture_dat <- read_excel("data/2_vulture_trap_deployments.xlsx")

# Import passcodes
MOVE_PASS <- Sys.getenv("MOVEBANK_PASSWORD")
MOVE_USE  <- Sys.getenv("MOVEBANK_USERNAME")

loginStored <- movebankLogin(username=MOVE_USE, 
                             password=MOVE_PASS)

# Get animals to check ID's in data
# Vultures
animals <-getMovebankAnimals(study=1573471517,login=loginStored)
# Ocelot
tmp <-getMovebankAnimals(study=2526574641,login=loginStored)
# carnivores
tmp1 <-getMovebankAnimals(study=2957292408,login=loginStored)
# Tapir
tmp2 <- getMovebankAnimals(study=1954804459,login=loginStored)
# Ecuador vultures
tmp3 <- getMovebankAnimals(study=2885579493,login=loginStored)
tmp3$local_identifier <- sub(".*? ", "", tmp3$local_identifier)
animals <- rbind(animals, tmp,tmp1, tmp2, tmp3)

# For some reason they are duplicated
animals[duplicated(animals)==F,]
# They vary by the field "sensor_type_id"
animals <- animals[animals$sensor_type_id==653 & is.na(animals$sensor_type_id)==F,]

# Clean up the name
animals$animalName <- paste0(sub('\\_.*', '', animals$animalName), "_", sub('\\ .*', '', animals$taxon_canonical_name))
animals$name <- sub('\\_.*', '', animals$animalName)


# Sort date objects
animals$timestamp_start <- ymd_hms(animals$timestamp_start)
animals$timestamp_end <- ymd_hms(animals$timestamp_end)
 

#drive_auth()
## Authenticate into googledrive service account ----
## 'GOOGLE_APPLICATION_CREDENTIALS' is what we named the Github Secret that 
## contains the credential JSON file
googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
data_folder <- drive_ls(path = "08 Databases + Analysis")
data_path <- "data"
dir.create(data_path) 
```


```{r, include=F}
#Capture_database
drive_download(file = as_id("1Q8ir4HNvP2zBFwItINhc1rmBvW2MdHnJ-mUULh_Z_oc"), path = paste0("data/cap_database"), overwrite = TRUE)
cap_db <- read_excel("data/cap_database.xlsx", sheet="database")  
# Convert the times to minutes
cap_db$total_time_processing <- as.numeric(cap_db$total_time_processing)*24*60
cap_db$delay <- as.numeric(cap_db$delay)*24*60
# Replace VALUE with NA
cap_db$individual_weight[cap_db$individual_weight=="#VALUE!"] <- NA
cap_db$individual_weight <- as.numeric(cap_db$individual_weight)

# Head
cap_db$head_length<- as.numeric(cap_db$head_length)
cap_db$head_width <- as.numeric(cap_db$head_width)

# Bill
cap_db$bill_length <- as.numeric(cap_db$bill_length)
cap_db$bill_width  <- as.numeric(cap_db$bill_width)
cap_db$bill_depth  <- as.numeric(cap_db$bill_depth)

cap_db$wing_chord  <- as.numeric(cap_db$wing_chord)
cap_db$wing_flat   <- as.numeric(cap_db$wing_flat)
cap_db$wing_span   <- as.numeric(cap_db$wing_span)
cap_db$tail_length <- as.numeric(cap_db$tail_length)

cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
cap_db$body_length <- as.numeric(cap_db$body_length)
cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

# Sort bological samples
cap_db$biological_sample <- as.numeric(cap_db$biological_sample)
# Setup the categories
cap_db$species <- as.factor(cap_db$species)


# Remove empty rows
cap_db <- cap_db[is.na(cap_db$date)==F & is.na(cap_db$country)==F,]

unique_db <- cap_db[,c("pit_tag_id", "local_identifier", "code_sp")]
# Remove NA's
unique_db <- unique_db[unique_db$local_identifier!="NA",]
# Remove duplicates
unique_db <- unique_db[duplicated(unique_db)==F,]
table(unique_db$local_identifier)

#trap_deployments
###################################
## THIS MAY NOT BE THE RIGHT ONE ##

drive_download(file = as_id("14hLi94W4WbDn6jdXahJMnxllOgLpZh-PpIBOWqt9jlc"), path = paste0("data/trap_database"), overwrite = TRUE)
trap_db1 <- read_excel("data/trap_database.xlsx")
trap_db1 <- trap_db1[is.na(trap_db1$date)==F,]

trap_db2 <- read_excel("data/trap_database.xlsx",sheet="Vulture_trap_deployments_Peru")
trap_db2 <- trap_db2[is.na(trap_db2$date)==F,]

# Delete when corrected
trap_db1$date <- as.character(trap_db1$date)
trap_db2$date <- as.character(trap_db2$date)

trap_db <- rbind(trap_db1, trap_db2)
remove(trap_db1, trap_db2)
trap_db <- trap_db[trap_db$deployment_id!="_",]
# Add a country code
trap_db$country <- NA
trap_db$country[trap_db$latitude>0] <- "CR"
trap_db$country[trap_db$latitude<0] <- "PE"


#biological samples
drive_download(file = as_id("1kx5BYcbbDYZLnWMp1IxQySaLRQ2I4AJhTNjZfn7m4U8"), path = paste0("data/bio_database"), overwrite = TRUE)
bio_db <- read_excel("data/bio_database.xlsx", sheet="biological_samples")
# Remove NA's
bio_db <- bio_db[is.na(bio_db$capture_date)==F,]
 
# Laboratory results
drive_download(file = as_id("1GYcqjqLG2SUKMF_s4f7JMGtgOt6snzEzHtFCsiwFiiQ"), path = paste0("data/lab_database"), overwrite = TRUE)
lab_db <- read_excel("data/lab_database.xlsx", sheet="Results")


# Laboratory results
drive_download(file = as_id("1GYcqjqLG2SUKMF_s4f7JMGtgOt6snzEzHtFCsiwFiiQ"), path = paste0("data/lab_database"), overwrite = TRUE)
lab_db <- read_excel("data/lab_database.xlsx", sheet="Results")


#device_database
drive_download(file = as_id("1_oa67Po8Hpmp4VH8jqlD_Sm25Fn2GPctqp7rC5EXSuo"), path = paste0("data/track_database"), overwrite = TRUE)
track_db <- read_excel("data/track_database.xlsx")

# Make one that is just the first of each unique code
track_db_1 <- track_db[!duplicated(track_db$transmitor_identifier) & is.na(track_db$deployment_start)==F,]


#pittag_database
drive_download(file = as_id("1oDWd3QYUxpSFRKkkwNLgRJmMJrxuMwqNfAM8tkN-UII"), path = paste0("data/pit_database"), overwrite = TRUE)
pit_db <- read_excel("data/pit_database.xlsx")
colnames(pit_db)[5:8] <- pit_db[1,5:8]
pit_db <- pit_db[is.na(pit_db$pit_tag_id)==F,]
# Remove blanks
pit_db <- pit_db[,1:9]

#cluster observations
drive_download(file = as_id("1KvZ2jEfv97eAJLatlw-rEic-kPitFPiUvjXcrrX7jiQ"), path = paste0("data/clus_database"), overwrite = TRUE)
clus_db <- read_excel("data/clus_database.xlsx", sheet="Database")


# Get current time and last 7 days
t <- now("America/Costa_Rica")
start_t <- t-as.difftime(7,units='days')

```

### Database changes (last 7 days)

The following table shows total number of observations in each database (left) and the new records from the last 7 days (right). The green numbers show where records have been added.

*Note* the changes are from when the data were collected, not when they were entered in Google Drive. If they were entered a long time after the collection date they may not be shown here. 

```{r, include=F}
cap_db$NEW   <- difftime(cap_db$date,t)>-7
trap_db$NEW  <- difftime(trap_db$date,t)>-7
bio_db$NEW   <- difftime(bio_db$capture_date,t)>-7
track_db_1$NEW <- difftime(track_db_1$deployment_start,t)>-7
clus_db$NEW   <- difftime(clus_db$visit_date,t)>-7

list_db <- list(cap_db, trap_db, bio_db, track_db_1, clus_db)
names(list_db) <- c("vulture_capture_database",
                                    "vulture_trap_deployments",
                                    "biological_samples",
                                    "tracking_device_deployment_database",
                                    "cluster_observations_database")


# Blank dataframe to store the output
change_dat <- data.frame(name=names(list_db),
                         total_rows = NA,
                         new_rows = NA,
                         CR_total = NA,
                         CR_new = NA,
                         PE_total =NA,
                         PE_new=NA)
i <- 2
for(i in 1:length(names(list_db)))
{
   #Overall
   tmp <- list_db[[i]] %>% group_by(NEW) %>% summarize(n=n())
   change_dat$total_rows[i] <- sum(tmp$n)
   # If you have new rows
   if(nrow(tmp)>1)
   {
     change_dat$new_rows[i] <- tmp$n[2]
   } else(change_dat$new_rows[i] <- 0)
   #Country specific
   tmp <- list_db[[i]] %>% group_by(NEW, country) %>% summarize(n=n())
   # Filter any NA;s
   tmp<- tmp[is.na(tmp$country)==F,]
   # CR
   tmp_cr <- tmp[tmp$country=="CR",]
   change_dat$CR_total[i] <- sum(tmp_cr$n)
   # If you have new rows
   if(nrow(tmp_cr)>1)
   {
     change_dat$CR_new[i] <- tmp_cr$n[2]
   } else(change_dat$CR_new[i] <- 0)
   # PE
   tmp_pe <- tmp[tmp$country=="PE",]
   change_dat$PE_total[i] <- sum(tmp_pe$n)
   # If you have new rows
   if(nrow(tmp_pe)>1)
   {
     change_dat$PE_new[i] <- tmp_pe$n[2]
   } else(change_dat$PE_new[i] <- 0)

  
}


```

```{r}
table_dat <- change_dat

# Setup highlights

table_dat$new_rows = cell_spec(table_dat$new_rows, background = ifelse(table_dat$new_rows == 0, "white", "lightgreen"))
table_dat$CR_new = cell_spec(table_dat$CR_new, background = ifelse(table_dat$CR_new == 0, "white", "lightgreen"))
table_dat$PE_new = cell_spec(table_dat$PE_new, background = ifelse(table_dat$PE_new == 0, "white", "lightgreen"))


# Setup column names 
colnames(table_dat)[2:7] <- c("total", "new", "total", "new", "total", "new") 

kbl(table_dat, escape = F) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, "All locations" = 2, "Costa Rica" = 2, "Peru" = 2)) %>% 
  column_spec (c(1,3,5,7),border_left = F, border_right = T) 
```

### Database linkage checks

There are several important checks to ensure our data integrity - each data entry sheet links to others (e.g. capture data to biological samples).

**Do all vulture captures have corresponding trap deployment data?**

```{r}
trap_db$date <- ymd(trap_db$date)

tmp <- cap_db[cap_db$date %in% trap_db$date,]
tmp2 <- cap_db[!(cap_db$date %in% trap_db$date),]

tmp3 <- data.frame("yes" = nrow(tmp), "no"=nrow(tmp2))

tmp3$no <- as.numeric(tmp3$no)

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)

```

The dates and locations when we captured vultures which are not in the database are as follows:

```{r}
tmp3 <- tmp2 %>% group_by(location_code, date) %>% summarize("vultures captured"=n())

kbl(tmp3, escape=F)  %>%
  kable_classic(full_width = F)
```

&nbsp;

**Do all captures with biological samples have data in the `biological_samples` database?**

```{r}
tmp <- cap_db[cap_db$biological_sample>0 & is.na(cap_db$biological_sample)==F,]
tmp2 <- bio_db


# Make a key variable
tmp$link <- paste(tmp$pit_tag_id, tmp$date, sep="_")
tmp2$link <- paste(tmp2$pit_tag_id, tmp2$capture_date, sep="_")

tmp3 <- data.frame("yes" = nrow(tmp[tmp$link %in% tmp2$link,]), "no"=nrow(tmp[!(tmp$link %in% tmp2$link),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

tmp <- tmp[!(tmp$link %in% tmp2$link),]

```

The captures whose biological samples are not currently in the database are:

```{r}
kbl(tmp[,c("local_identifier", "pit_tag_id", "date")])  %>%
  kable_classic(full_width = F)
```
&nbsp;

**Do any biological samples have duplicated codes?**

```{r}
tmp3 <- bio_db[duplicated(bio_db$sample_code)==T,]

tmp2 <- data.frame("no" = nrow(tmp2)-nrow(tmp3), "yes"=nrow(tmp3))

kbl(tmp2)  %>%
  kable_classic(full_width = F)

tmp2 <- bio_db[bio_db$sample_code %in% tmp3$sample_code, ]
tmp2 <- tmp2[order(tmp2$sample_code, tmp2$pit_tag_id),]

```

If there are any present, the samples with duplicated `sample_codes` are shown below:

```{r}
if(nrow(tmp2)>0){
kbl(tmp2[,colnames(tmp2)!="specific_id"])  %>%
  kable_classic(full_width = F)
}else{print("No samples duplicated")}
```

&nbsp;

**Do all vulture captures where a device was deployed have data in `device_deployment_database`?**

```{r}
cap_db$transmitor_installation <- as.numeric(cap_db$transmitor_installation)
tmp <- cap_db[cap_db$transmitor_installation==1 & is.na(cap_db$transmitor_installation)==F,]
tmp2 <- track_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$transmitor_identifier %in% tmp2$transmitor_identifier,]), "no"=nrow(tmp[!(tmp$transmitor_identifier %in% tmp2$transmitor_identifier),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)

```
&nbsp;

**Do all vultures with pit-tags have an entry in the `vulture_pit_tag` database?**

```{r}
tmp <- cap_db[nchar(cap_db$pit_tag_id)>10,]
tmp2 <- pit_db


tmp3 <- data.frame("yes" = nrow(tmp[tmp$pit_tag_id %in% tmp2$pit_tag_id,]), 
                   "no"  = nrow(tmp[!(tmp$pit_tag_id %in% tmp2$pit_tag_id),]))

kbl(tmp3)  %>%
  kable_classic(full_width = F)
```
&nbsp;

**Are all vultures in the `device_deployment_database` represented in Movebank?**

And if any are not, what are their names?

```{r}
tmp <- unique(track_db$local_identifier) 
tmp <- tmp[is.na(tmp)==F] 
tmp2 <- table(tmp %in% animals$local_identifier)

kbl(tmp2)  %>%
  kable_classic(full_width = F)

print(tmp[!(tmp %in% animals$local_identifier)])

```
&nbsp;

### Recaptures

To date we have had 
`r nrow(cap_db[duplicated(cap_db$pit_tag_id)==T,])` vulture recaptures. The ID's and dates are as follows:

```{r}

tmp <- cap_db[duplicated(cap_db$pit_tag_id)==T,]
tmp2 <- cap_db[cap_db$pit_tag_id %in% tmp$pit_tag_id,]
tmp2 <- tmp2[order(tmp2$pit_tag_id, tmp2$date),]
tmp <- tmp2[, c("local_identifier", "species",  "pit_tag_id", "date", "location_code")]

kbl(tmp)  %>%
  kable_classic(full_width = F)
```

Are these expected?


# Capture database

To date we have captured vultures `r nrow(cap_db)` times, deployed `r length(unique(cap_db$transmitor_identifier))` tracking devices and `r length(unique(cap_db$pit_tag_id[nchar(cap_db$pit_tag_id)>10]))` pit tags. 

These captures occurred on the following dates:

```{r}

tmp <- cap_db
tmp$country <- as.factor(tmp$country)
tmp$location <- as.numeric(tmp$country)
tmp$local_identifier[tmp$local_identifier=="NA"] <- tmp$pit_tag_id[tmp$local_identifier=="NA"] 
plot_ly(data = tmp, x = tmp$date, y=tmp$location, color = ~species,
               hoverinfo = 'text',height=200,
               text = ~paste(date,local_identifier,sep = "<br>")) %>% 
              layout(title = '', 
                      xaxis = list(title = 'Vulture capture date'), 
                      yaxis = list(title = "",
                                   ticktext=c("Costa Rica", "Peru"),
                                   tickvals=list(1,2)), 
                      legend = list(title=list(text='<b> Species </b>')))

```


The following section provides a summary of the data taken during animal captures

### Processing time

Our average processing times for the different species are as follows:

```{r}
tmp <-cap_db %>% group_by(species) %>% summarise("Mean processing time (mins)" = round(mean(total_time_processing, na.rm=T),0))

kbl(tmp)  %>%
  kable_classic(full_width = F)

```



```{r}
# Sensors deploy and sensorys not deployed

#plot(cap_db$total_time_processing, ylab="Processing time (mins)")

fig <- plot_ly(data = cap_db, x = 1:nrow(cap_db), y = ~total_time_processing, color = ~species,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,total_time_processing,sep = "<br>")) %>% 
              layout(title = 'Vulture processing time', 
                      xaxis = list(title = 'Vulture capture order'), 
                      yaxis = list(title = 'Processing time (minutes)'), 
                      legend = list(title=list(text='<b> Species </b>')))




fig

```

### Crop size

```{r}
cap_db$crop <- as.numeric(cap_db$crop)
tmp <- as.data.frame.matrix(table(cap_db$species,cap_db$crop))

kbl(tmp) %>%
  kable_classic(full_width = F) %>%   add_header_above(header = c("Species" = 1, "crop" = 3))

```

### Morphometrics

*Weight* 

Of `r nrow(cap_db)` captures we have *`r length(cap_db$individual_weight[is.na(cap_db$individual_weight)==T])`* captures where weight was not assessed.

Of the weights recorded the distributions are as follows:

```{r}
# Weight

f1 <- plot_ly(data=cap_db, y = ~individual_weight, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,individual_weight,sep = "<br>")) %>% 
              layout(title = 'Individual weight (g)', 
                      xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Weight (g)'))

f1


```


Other:

**Head traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~head_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,head_length,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~head_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,head_width,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.2, y = 1.0,   text = "Head length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.8, y = 1,text = "Head width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

**Bill traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~bill_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,bill_length,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~bill_width, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,bill_width,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~bill_depth, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,bill_depth,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Bill length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Bill width (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Bill depth (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig






```

**Wing traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~wing_chord, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,wing_chord,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~wing_flat, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,wing_flat,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~wing_span, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,wing_span,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)

f4 <- plot_ly(data=cap_db, y = ~tail_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,tail_length,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = ''),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Wing chord (cm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.35, y = 1,text = "Wing flat (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.65, y = 1,text = "Wing span (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Tail length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3, f4) 
fig <- fig %>%layout(annotations = annotations) 
fig


# cap_db$tarsus_length <- as.numeric(cap_db$tarsus_length)
# cap_db$body_length <- as.numeric(cap_db$body_length)
# cap_db$MTO_length <- as.numeric(cap_db$MTO_length)

```

**Size traits**

```{r}
f1 <- plot_ly(data=cap_db, y = ~tarsus_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,tarsus_length,sep = "<br>")) %>% 
              layout( xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f2 <- plot_ly(data=cap_db, y = ~body_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,body_length,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)

f3 <- plot_ly(data=cap_db, y = ~MTO_length, color = ~species, 
               type = "box", boxpoints = "all", jitter = 0.3, pointpos = -1.8,
               hoverinfo = 'text',
               text = ~paste(date,pit_tag_id,MTO_length,sep = "<br>")) %>% 
              layout(xaxis = list(title = 'Species'), 
                      yaxis = list(title = 'Head length (mm)'),
                     showlegend=F)


annotations = list( 
                    list(x = 0.1, y = 1.0,   text = "Tarsus length (mm)",
                         xref = "paper",  yref = "paper",  xanchor = "center", yanchor = "bottom", 
                         showarrow = FALSE),
                    list(x = 0.5, y = 1,text = "Body length (cm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE),
                    list(x = 0.9, y = 1,text = "Middle toe length (mm)",
                         xref = "paper",yref = "paper", xanchor = "center", yanchor = "bottom",  
                         showarrow = FALSE)
                  )

fig <- subplot(f1, f2, f3) 
fig <- fig %>%layout(annotations = annotations) 
fig
```

# Trap deployment database

To date we have deployed vulture traps in `r  length(unique(trap_db$location_code)) ` different locations.

The trapping locations are as follows:

```{r}
tmp <- trap_db[,c("location_code", "latitude", "longitude")]
tmp <- tmp[duplicated(tmp)==F,]


kbl(tmp)  %>%
  kable_classic(full_width = F)


```


```{r}

m <- leaflet() %>%
      fitBounds(-70,-12.8,-84,8.8) %>% 
      # Add a satellite image layer
      addProviderTiles(providers$CartoDB.PositronNoLabels, group="Simple") %>%
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=tmp$longitude,
                         lat=tmp$latitude, 
                         popup=paste(tmp$location_code)) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Simple", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 

```

# Biological samples database

To date we have collected `r nrow(bio_db)` biological samples from `r length(unique(bio_db$pit_tag_id))` individuals. The species breakdown is as follows:

```{r}
# add a species column
bio_db$species <- NA
#cap_db$species
bio_db$species[bio_db$code_sp=="BV"] <- "coragyps_atratus"
bio_db$species[bio_db$code_sp=="KV"] <- "sarcoramphus_papa"
bio_db$species[bio_db$code_sp=="LP"] <- "leopards_pardalis"
bio_db$species[bio_db$code_sp=="TV"] <- "cathartes_aura"
bio_db$species[bio_db$code_sp=="GYV"] <- "cathartes_melambrotus"
bio_db$species[bio_db$code_sp=="TB"] <- "tapirus_bairdii"


tmp <- bio_db %>% group_by(species) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F)

```
 
Break down of the type of sample per species:

```{r}
tmp <- bio_db %>% group_by(species, tissue ) %>% summarise(individuals=n_distinct(pit_tag_id), samples=n())

kbl(tmp[, colnames(tmp)!="species"])  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp$species))

```

And the current locations of the samples:

```{r}
tmp <- bio_db %>% group_by(tissue, storage_box) %>% summarise(samples=n())

kbl(tmp)  %>%
  kable_classic(full_width = F) 

```

# Tracking device deployment database

To date we have deployed `r nrow(track_db_1)` tracking devices.

### Do all the current vital settings match?

It is important that all of our common tag types have the same settings. Is this currently the case in the database?

We first subset to the most recent deployment settings for each tag, then compare them the most important categories:

```{r}
tmp <- track_db[is.na(track_db$deployment_start)==F,]
tmp <- tmp[order(tmp$local_identifier,tmp$deployment_start, decreasing=T),]

tmp_latest <- tmp[duplicated(tmp$local_identifier)==F,]

# Convert times into something sensible
tmp_latest$`ACC On From (HH:MM)` <- round((as.numeric(tmp_latest$`ACC On From (HH:MM)`)*24),2)
tmp_latest$`ACC On Until (HH:MM)`<- round((as.numeric(tmp_latest$`ACC On Until (HH:MM)`)*24),2)

tmp_latest$`GPS On From (HH:MM)` <- substr(tmp_latest$`GPS On From (HH:MM)`,12,16)
tmp_latest$`GPS On Until (HH:MM)`<- substr(tmp_latest$`GPS On Until (HH:MM)`,12,16)

tmp_latest$`Pinger On From (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On From (HH:MM)`)*24),2)

tmp_latest$`Pinger On Until (HH:MM)` <- round((as.numeric(tmp_latest$`Pinger On Until (HH:MM)`)*24),2)


tmp2 <- tmp_latest[, c("country", "transmitor_identifier", "local_identifier", 
               "tag_model", "ACC On From (HH:MM)", "ACC On Until (HH:MM)",
               "ACC Interval (s)",
               "GPS On From (HH:MM)", "GPS On Until (HH:MM)",
               "GPS LOW RES INTERVAL (s)", "GPS HIGH RES INTERVAL (s)",
               "Pinger On From (HH:MM)", "Pinger On Until (HH:MM)",
               "Freq. (MHz)",
               "GPRS Time (HH:MM)", "GPRS Time 2 (HH:MM)"),
           ]

# Order by country and tag model
tmp2 <- tmp2[order(tmp2$country, tmp2$tag_model, tmp2$local_identifier),]

kbl(tmp2)  %>%
  kable_classic(full_width = F) %>%
  pack_rows(index=table(tmp2$country))%>%
    scroll_box(width = "100%", height="600px")


```

# Laboratory analyses

We first need to check if all biological samples have associated pit-tag codes, and then add in the `local_identifier` information, if it is available.

**Do all biological samples have associated pit-tag codes?**

```{r}
# Add in the pittag code
lab_full <- left_join(lab_db, bio_db[,c("sample_code", "pit_tag_id", "code_sp")])
# And if individuals have a unique ID - add it
lab_full <- left_join(lab_full, unique_db)

kbl(table(is.na(lab_full$pit_tag_id)==F)) %>%
  kable_classic(full_width = F)

```

To date we have processed samples with the following methodologies:

```{r}

tmp <- lab_full %>% group_by(assay) %>% summarise(total_assays=n(), individuals=n_distinct(pit_tag_id))

kbl(tmp) %>%
  kable_classic(full_width = F)

```

### Sex

So far we have sexed `r nrow(lab_full[lab_full$assay=="sex",])` vultures:

```{r}
tmp <- lab_full[lab_full$assay=="sex",]
tmp <- tmp[,c("local_identifier", "result", "code_sp")]
tmp <- tmp[order(tmp$code_sp,tmp$result, tmp$local_identifier),]
kbl(tmp) %>%
  kable_classic(full_width = F) 
```

### Blood cell counts

```{r}
tmp <- lab_full[lab_full$assay=="blood",]

# Add in country data (and individual)
#tmp2 <- cap_db[,c("local_identifier", "pit_tag_id", "country", "specific_id", "date")]
#colnames(tmp2)[5] <- "capture_date"
#tmp3 <- bio_db[, c("specific_id", "capture_date", "sample_code")]


#id_loc_dat <- left_join(tmp3, tmp2, by=c("specific_id", "capture_date"))
# Add in the ID and location data

#tmp
loc_code <- cap_db[,c("pit_tag_id", "country")]
# Remove recaptures
loc_code <- loc_code[duplicated(loc_code)==F,]
tmp <- left_join(tmp, loc_code)


```

We have blood cell parameters quantified for `r length(unique(tmp$pit_tag_id))` individuals and `r length(unique(tmp$code_sp))` species. 

```{r}
tmp2 <- tmp %>% group_by(country, code_sp) %>% summarise(individuals=n_distinct(pit_tag_id))

kbl(tmp2) %>%
  kable_classic(full_width = F) 
```

Differentials first:

```{r}
#table(tmp$units)
tmp$result <- as.numeric(tmp$result)


tmp2 <- tmp[tmp$assay_target %in% c("basophils", "eosinophils", "Hematocrit", "heterophils", "lymphocytes", "monocytes"),]
tmp2 <- tmp2[tmp2$units=="%",]
#length(unique(dat$local_identifier))
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
# Make the local identifier name blank is NA
tmp2$local_identifier[is.na(tmp2$local_identifier)==T] <- tmp2$pit_tag_id[is.na(tmp2$local_identifier)==T]

params <- tmp2$assay_target

fig <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp, 
               type = "box", boxpoints = "all", jitter = 0.1,
        pointpos = 0, hovertext=paste(tmp2$local_identifier)) %>% 
  layout(yaxis = list(title = '"% composition"'), boxmode="group")

fig

```

And the other parameters:

```{r}
dat <- tmp[tmp$method=="Cell Count" & is.na(tmp$method)==F & tmp$units!="%" & !(tmp$assay_target %in% params),]
dat$local_identifier[is.na(dat$local_identifier)==T] <- dat$pit_tag_id[is.na(dat$local_identifier)==T]
dat$result <- as.numeric(dat$result)
#boxplot(dat$result~dat$assay_target, las=1, ylab=dat$units[1],
#        xlab="Contaminant")
tmp2 <- dat[dat$assay_target%in% c("Erythrocytes","proportion_H:L", "thrombocytes", "leucocytes", "Hemoglobin") ,]

fig1 <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp,
                 type = "box", boxpoints = "all", jitter = 0.2,
        pointpos = 0, hovertext=tmp2$local_identifier) %>% 
  layout(yaxis = list(title = "units", rangemode="tozero"), boxmode="group")

fig1

tmp2 <- dat[!(dat$assay_target%in% c("Erythrocytes","proportion_H:L", "thrombocytes", "leucocytes", "Hemoglobin")) ,]

fig1 <- plot_ly(y =tmp2$result,x=tmp2$assay_target, color=tmp2$code_sp,
                 type = "box", boxpoints = "all", jitter = 0.2,
        pointpos = 0, hovertext=tmp2$local_identifier) %>% 
  layout(yaxis = list(title = "units", rangemode="tozero"), boxmode="group")

fig1
```

### Summary tables

```{r}
tmp$units[tmp$units=="NA"] <- ""
tmp$name <- paste0(tmp$assay_target, " (", tmp$units,")")

res_sum <- tmp %>% group_by(code_sp, name) %>% summarise(mean=mean(result), sd=sd(result), min=min(result), max=max(result))
res_sum <- data.frame(res_sum)

colnames(res_sum)[2] <- "Test"
# Sort the number of dp
res_sum$mean <- round(res_sum$mean,2)
res_sum$sd <- round(res_sum$sd,1)
res_sum$min <- round(res_sum$min,2)
res_sum$max <- round(res_sum$max,2)


```

King vulture (`r length(unique(tmp$sample_code[tmp$code_sp=="KV"]))` samples):

```{r}
kbl(res_sum[res_sum$code_sp=="KV",2:ncol(res_sum)], escape = F,row.names = FALSE) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, 
                     "King vulture" =4))

```


Turkey vulture (`r length(unique(tmp$sample_code[tmp$code_sp=="TV"]))` samples):

```{r}
kbl(res_sum[res_sum$code_sp=="TV",2:ncol(res_sum)], escape = F,row.names = FALSE) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, 
                     "Turkey vulture" =4))

```


Black vulture (`r length(unique(tmp$sample_code[tmp$code_sp=="BV"]))` samples):

```{r}
kbl(res_sum[res_sum$code_sp=="BV",2:ncol(res_sum)], escape = F,row.names = FALSE) %>%
  kable_classic(full_width = T) %>%
  add_header_above(c(" " = 1, 
                     "Black vulture" =4))

```



# Cluster visits

To date we have visited `r nrow(clus_db)' cluster sites/tracking locations.

The visits dates are as follows, each point represents a visit:

```{r}
clus_db$country <- as.factor(clus_db$country)
clus_db$location <- as.numeric(clus_db$country)

plot_ly(data = clus_db, x = clus_db$visit_date, y=clus_db$location,
               hoverinfo = 'text',
               text = ~paste(visit_date,animal_id,sep = "<br>"), height=200) %>% 
              layout(title = 'Cluster visits', 
                      xaxis = list(title = 'Dates'), 
                      yaxis = list(title = "",
                                   ticktext=c("Costa Rica", "Peru"),
                                   tickvals=list(1,2)))


```

Where are these clusters?

```{r}
# Define a palette for the colourscheme
col_pal <- "VanGogh2"
clus_db$site_type<- as.factor(clus_db$site_type)
col.cat <- as.character(met.brewer(col_pal,n=length(levels(clus_db$site_type))))

clus_db$Cols <- col.cat[clus_db$site_type]

m <- leaflet() %>%
      # Add a satellite image layer
      addProviderTiles(providers$CartoDB.PositronNoLabels, group="Simple") %>%
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=clus_db$cluster_longitude,
                         lat=clus_db$cluster_latitude, 
                         popup=paste(clus_db$visit_date, clus_db$animal_id,
                                     "<br>", clus_db$site_type),
                         color=clus_db$Cols,
                         fillColor = clus_db$Cols,
                          opacity=0.8) %>%
      addLegend("topright", colors = col.cat,  labels = levels(clus_db$site_type),
      title = "Site type",
      labFormat = labelFormat(prefix = "$"),
      opacity = 1) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Simple", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 
```

Of the clusters visited so far, the type of carrion found was as follows (note NA's = no carcass).

```{r}
kbl(table(clus_db$carcass_id))%>%
  kable_classic(full_width = F) 
```

On how many occasions have we observed/detected a animal with a GPS device?

```{r}
kbl(table(clus_db$focal_animal_detected))%>%
  kable_classic(full_width = F) 

```

Who were the animals observed?

```{r}
kbl(table(clus_db$animal_id[clus_db$focal_animal_detected=="yes"]))%>%
  kable_classic(full_width = F) 
```


# Exclosure data

```{r}
# Import the exclosure data - it is stored in two locations:

# 1. https://docs.google.com/spreadsheets/d/1i6Isj2vmtzdljYw6h2D29pdqLYxByFI_XsiB-Oy8lZE/edit?usp=sharing 

drive_download(file = as_id("1i6Isj2vmtzdljYw6h2D29pdqLYxByFI_XsiB-Oy8lZE"), path = paste0("data/exc_database"), overwrite = TRUE)
ex_check_db <- read_excel("data/exc_database.xlsx", sheet="Carcass_checks")

# Cleaning 
  # Remove NA's
  ex_check_db <- ex_check_db[is.na(ex_check_db$country)==F,]

  # Remove the trial data
  ex_check_db <- ex_check_db[ex_check_db$round!="trial",]
  # Convert to numeric
  ex_check_db$round <- as.numeric(ex_check_db$round)

  # Create a unique column by pasting the location_code and round 
  ex_check_db$replicate  <- paste(ex_check_db$treatment,ex_check_db$habitat, ex_check_db$round, sep="_")

  # Create a days column of days since deployment
  ex_check_db$days <- difftime(ex_check_db$visit_date, ex_check_db$date_carcass_deployed)/ddays(1) 
  
  
  #table(ex_check_db$replicate)
ex_locs <- read_excel("data/exc_database.xlsx", sheet="Locations_masterlist")



# The species data
# 2. https://docs.google.com/spreadsheets/d/18WlT7Loq0xBST6vR8vqt333whm6wJJBCSPyWaz5Z6M0/

drive_download(file = as_id("18WlT7Loq0xBST6vR8vqt333whm6wJJBCSPyWaz5Z6M0"), path = paste0("data/exc_samples"), overwrite = TRUE)
ex_db <- read_excel("data/exc_samples.xlsx", sheet="DATA")
ex_sp <- read_excel("data/exc_samples.xlsx", sheet="SPECIES_CHARACT")

# Remove NAs
ex_db <- ex_db[is.na(ex_db$date_carcass_deployed)==F,]


 # Create a unique column by pasting the location_code and round
  ex_db$replicate  <- paste(ex_db$treatment,ex_db$habitat, ex_db$round, sep="_")

  # Create a days column of days since deployment
  ex_db$days <- difftime(ex_db$sample_date, ex_db$date_carcass_deployed)/ddays(1)
#   

```

### Map of sampling locations

```{r}
# Create a plot of survey locations
ex_locs$Habitat_type <- as.factor(ex_locs$Habitat_type)
cols <- c("#e41a1c", "#ff7f00")
m <- leaflet() %>% 
      # Add a satellite image layer
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=ex_locs$Longitude,
                         lat=ex_locs$Latitude, 
                         popup=paste(ex_locs$Location_code, ex_locs$Habitat_type),
                        color=cols[ex_locs$Habitat_type],
                       opacity=0.8) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) %>% 
  addLegend(colors = cols, labels=c("Closed" ,"Open"), opacity = 1)
  
m 
```

### Carcass weights through time

Absolute:

```{r}
ex_check_db$real_current_weight_kg <- as.numeric(ex_check_db$real_current_weight_kg)
ex_check_db$full_treat <- as.factor(paste(ex_check_db$treatment, ex_check_db$habitat, sep="_"))
# Order the dataframe
ex_check_db <- ex_check_db[order(ex_check_db$round,ex_check_db$full_treat,ex_check_db$days),]
tmp <- unique(ex_check_db$replicate)


cols<-c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")


fig <- plot_ly() 
for(i in 1:length(tmp))
{
   fig <- fig %>% add_trace(x = ex_check_db$days[ex_check_db$replicate==tmp[i]], y=ex_check_db$real_current_weight_kg[ex_check_db$replicate==tmp[i]],
                            mode = 'lines+markers', type="scatter",legendgroup=ex_check_db$round[ex_check_db$replicate==tmp[i]][1],
                            name=ex_check_db$replicate[ex_check_db$replicate==tmp[i]][1],  connectgaps = TRUE, 
                            color=cols[ex_check_db$full_treat[ex_check_db$replicate==tmp[i]][1]]) 
}
fig %>% 
  layout(title= list(text = "Absolute change"), font=t, 
         legend=list(title=list(text='Treatment')), 
         xaxis = list(title = list(text ='Time (days)')),
         yaxis = list(title = list(text ='Carcass remaining (kg)')))

```

Percentage:

```{r}
ex_check_db$real_current_weight_kg <- as.numeric(ex_check_db$real_current_weight_kg)

ex_check_db$percentage_weight <- round(ex_check_db$real_current_weight_kg/as.numeric(ex_check_db$initial_bait_weight_kg),2)*100

# Order the dataframe
ex_check_db <- ex_check_db[order(ex_check_db$round,ex_check_db$full_treat,ex_check_db$days),]
tmp <- unique(ex_check_db$replicate)


cols<-c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")
ex_check_db$full_treat <- as.factor(paste(ex_check_db$treatment, ex_check_db$habitat, sep="_"))

fig <- plot_ly() 
for(i in 1:length(tmp))
{
   fig <- fig %>% add_trace(x = ex_check_db$days[ex_check_db$replicate==tmp[i]], y=ex_check_db$percentage_weight[ex_check_db$replicate==tmp[i]],
                            mode = 'lines+markers', type="scatter",legendgroup=ex_check_db$round[ex_check_db$replicate==tmp[i]][1],
                            name=ex_check_db$replicate[ex_check_db$replicate==tmp[i]][1],  connectgaps = TRUE, 
                            color=cols[ex_check_db$full_treat[ex_check_db$replicate==tmp[i]][1]]) 
}
fig %>% 
  layout(title= list(text = "Percentage change"), font=t, 
         legend=list(title=list(text='Treatment')), 
         xaxis = list(title = list(text ='Time (days)')),
         yaxis = list(title = list(text ='Carcass remaining (%)')))

```


### Capture rates by treatment and method

```{r}
ex_db_long <- ex_db %>%
  pivot_longer(
    cols = starts_with("sp"),
    #names_to = "week",
    #names_prefix = "wk",
    values_to = "count",
    values_drop_na = TRUE
  )


cr_summary <- ex_db_long %>% group_by(replicate, trap_type, round, days) %>% summarize(total_ind=sum(count), total_sp=n_distinct(name))

cols<-c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")
i <- 1

# Create a way to store multiple plots

l <- htmltools::tagList()

for(i in 1:length(unique(cr_summary$round)))
{
  fig <- plot_ly( yaxis = list(type = "log"))
  tmp <- cr_summary[cr_summary$round==i,]
  treats <- unique(tmp$replicate)[order(unique(tmp$replicate))]
  methods <- unique(tmp$trap_type)[order(unique(tmp$trap_type))]
  for(j in 1:length(treats))
  {
       fig <- fig %>% add_trace(x = tmp$days[tmp$replicate==treats[j] & tmp$trap_type==methods[1]], 
                                y = tmp$total_ind[tmp$replicate==treats[j] & tmp$trap_type==methods[1]],
                                mode = 'lines+markers', type="scatter",
                                legendgroup = methods[1],
                                name=paste(methods[1],treats[j]),  
                                connectgaps = TRUE, 
                                color=cols[j]) 
       fig <- fig %>% add_trace(x = tmp$days[tmp$replicate==treats[j] & tmp$trap_type==methods[2]], 
                                y = tmp$total_ind[tmp$replicate==treats[j] & tmp$trap_type==methods[2]],
                                mode = 'lines+markers', type="scatter",
                                line = list(widthh=0.5, dash="dot"),
                                legendgroup = methods[2],
                                name=paste(methods[2],treats[j]),  
                                connectgaps = TRUE, 
                                color=cols[j]) 

  }
  l[[i]] <- as_widget(fig %>% 
  layout(title= list(text = paste("Round", i, "Individuals")), font=t, 
         legend=list(title=list(text='Method Treatment')), 
         xaxis = list(title = list(text ='Time (days)')),
         yaxis = list(title = list(text ='Total individuals'))))

}

l
```

### Morphospecies by treatment and method

```{r}

l <- htmltools::tagList()

for(i in 1:length(unique(cr_summary$round)))
{
  fig <- plot_ly( yaxis = list(type = "log"))
  tmp <- cr_summary[cr_summary$round==i,]
  treats <- unique(tmp$replicate)[order(unique(tmp$replicate))]
  methods <- unique(tmp$trap_type)[order(unique(tmp$trap_type))]
  for(j in 1:length(treats))
  {
       fig <- fig %>% add_trace(x = tmp$days[tmp$replicate==treats[j] & tmp$trap_type==methods[1]], 
                                y = tmp$total_sp[tmp$replicate==treats[j] & tmp$trap_type==methods[1]],
                                mode = 'lines+markers', type="scatter",
                                legendgroup = methods[1],
                                name=paste(methods[1],treats[j]),  
                                connectgaps = TRUE, 
                                color=cols[j]) 
       fig <- fig %>% add_trace(x = tmp$days[tmp$replicate==treats[j] & tmp$trap_type==methods[2]], 
                                y = tmp$total_sp[tmp$replicate==treats[j] & tmp$trap_type==methods[2]],
                                mode = 'lines+markers', type="scatter",
                                line = list(widthh=0.5, dash="dot"),
                                legendgroup = methods[2],
                                name=paste(methods[2],treats[j]),  
                                connectgaps = TRUE, 
                                color=cols[j]) 

  }
  l[[i]] <- as_widget(fig %>% 
  layout(title= list(text = paste("Round", i, "Morphospecies")), font=t, 
         legend=list(title=list(text='Method Treatment')), 
         xaxis = list(title = list(text ='Time (days)')),
         yaxis = list(title = list(text ='Total Morphospecies'))))
}

l


```



```{r}
### Species compositions
#library(tidyr)

#ex_sp





```


# Farm surveys

```{r}
# Import spreadsheet
drive_download(file = as_id("1vjlbdgLLHHzq-lLR2LOYiUr89471F_uJifnKDDAYrRA"), path = paste0("data/farm_database"), overwrite = TRUE)
farm_db <- read_excel("data/farm_database.xlsx", sheet="Standarize answer")  

# Sort coordinates
farm_db$long <- as.numeric(substr(farm_db$`Farms Location Gps coordinates`, 15,23))*-1 
farm_db$lat <- as.numeric(substr(farm_db$`Farms Location Gps coordinates`, 4,10))
```

To date we have completed `r nrow(farm_db)` surveys. The locations where surveys were conducted are:

```{r}
m <- leaflet() %>% 
      # Add a satellite image layer
      addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
      addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%     
      addCircleMarkers(lng=farm_db$long,
                         lat=farm_db$lat, 
                         popup=paste(farm_db$`Name of the farm`, farm_db$`Name of the farmer`),
                       opacity=0.8) %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("OS", "Satellite"),
        options = layersControlOptions(collapsed = FALSE)
      ) 
  
m 
```

The following plots summarise the responses:

### Are you a full time farmer?

```{r}
tmp <- farm_db %>% group_by(`Are you a full-time farmer based on the farm?`) %>% summarise(freq=n())


fig <- plot_ly(tmp, labels = ~`Are you a full-time farmer based on the farm?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### What is the size of the farm?

```{r}
fig <- plot_ly(x=farm_db$`What is the size of the farm in hectares?`, type = "histogram")

fig
```

Most are small (1 - 190 Ha).  

### Number of animals

```{r}

f1 <- plot_ly(x=as.numeric(farm_db$Cows), type = "histogram", name="Cows") %>% layout(title = 'Cows')
f2 <- plot_ly(x=as.numeric(farm_db$Pig), type = "histogram", name="Pigs") %>% layout(title = 'Pigs')
f3 <- plot_ly(x=as.numeric(farm_db$Chicken), type = "histogram", name="Chicken") %>% layout(title = 'Chicken')
f4 <- plot_ly(x=as.numeric(farm_db$Horses), type = "histogram", name="Horses") %>% layout(title = 'Horses')
f5 <- plot_ly(x=as.numeric(farm_db$Sheeps), type = "histogram", name="Sheep") %>% layout(title = 'Sheep')
f6 <- plot_ly(x=as.numeric(farm_db$Goats), type = "histogram", name="Goats") %>% layout(title = 'Goats')
fig <- subplot(f1, f2, f3, f4, f5, f6, nrows=2) %>% 
  layout(title = 'Animals on the farm')
fig
```


### Other crops

```{r}
table(c(farm_db$`Monoculture crops`, farm_db$`Polyculture crops`))
```

### Main production

```{r}
#farm_db$`What is your main production?`

tmp <- farm_db %>% group_by(`What is your main production?`) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What is your main production?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Do you vaccinate?

```{r}
#farm_db$`Do you vaccinate your animals?`

tmp <- farm_db %>% group_by(`Do you vaccinate your animals?`) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Do you vaccinate your animals?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### What do you feed your cattle?

```{r}
#farm_db$`What do your animals feed on additionnal to grazing?`

tmp <- farm_db %>% group_by(`What do your animals feed on additionnal to grazing?`) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What do your animals feed on additionnal to grazing?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Animal diseases

```{r}
#farm_db$`What are the diseases that affect your animals?`

tmp <- farm_db %>% group_by(`What are the diseases that affect your animals?`) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What are the diseases that affect your animals?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### How do you manage sick animals?

```{r}
#farm_db$`How do you manage sick animals?`

tmp <- farm_db %>% group_by(`How do you manage sick animals?`) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`How do you manage sick animals?`, values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Where do you source water for your animals?  

```{r}
#farm_db$`What is the water source use for the animals from?`     

tmp <- farm_db %>% group_by(`What is the water source use for the animals from?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What is the water source use for the animals from?`  , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Do you treat your animal when sick?  

```{r}
#farm_db$`Do you treat your animal when sick?`    

tmp <- farm_db %>% group_by(`Do you treat your animal when sick?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Do you treat your animal when sick?`  , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### What antibiotics do you use?  

```{r}
#farm_db$`What antibiotics do you use?`   

tmp <- farm_db %>% group_by(`What antibiotics do you use?`   ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What antibiotics do you use?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Have you experienced any disease outbreaks on your farm in the past 2 years? 

```{r, eval=F}
#farm_db$`Have you experienced any disease outbreaks on your farm in the past 2 years?`   

tmp <- farm_db %>% group_by(`Have you experienced any disease outbreaks on your farm in the past 2 years?`   ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Have you experienced any disease outbreaks on your farm in the past 2 years?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```


```{r}
#farm_db$`If yes, what diseases.`   

tmp <- farm_db %>% group_by(`If yes, what diseases.`   ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`If yes, what diseases.`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### How often do animals die? 

```{r}
#farm_db$`How often do animals die?`   

tmp <- farm_db %>% group_by(`How often do animals die?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`How often do animals die?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Normal Cause of death?

```{r}
#farm_db$`Normal Cause of death?`   

tmp <- farm_db %>% group_by(`Normal Cause of death?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Normal Cause of death?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### How often do you consult with a veterinarian and what is their role?

```{r}
#farm_db$`How often do you consult with a veterinarian and what is their role?`  

tmp <- farm_db %>% group_by(`How often do you consult with a veterinarian and what is their role?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`How often do you consult with a veterinarian and what is their role?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Do you consider veterinary services to be affordable and accessible?

```{r}
#farm_db$`Do you consider veterinary services to be affordable and accessible?`  

tmp <- farm_db %>% group_by(`Do you consider veterinary services to be affordable and accessible?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Do you consider veterinary services to be affordable and accessible?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Have you observed vultures?

#### Black

```{r}

tmp1 <- farm_db %>% group_by(`Have you observed black vultures in your farm? How often?`  ) %>% summarise(freq=n())
f1 <- plot_ly(tmp1, labels = ~`Have you observed black vultures in your farm? How often?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))
f1
```

#### King

```{r}
tmp2 <- farm_db %>% group_by( `Have you observed turkey vultures in your farm? How often?` ) %>% summarise(freq=n())
f2 <- plot_ly(tmp2, labels = ~`Have you observed turkey vultures in your farm? How often?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))
f2

```

#### Turkey

```{r}
tmp3 <- farm_db %>% group_by( `Have you observed king vultures in your farm? How often?` ) %>% summarise(freq=n())
f3 <- plot_ly(tmp3, labels = ~`Have you observed king vultures in your farm? How often?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))
f3
```

#### Lesser-yellow

```{r}
tmp4 <- farm_db %>% group_by( `Have you observed Lesser yellow headed vultures in your farm? How often?` ) %>% summarise(freq=n())
f4 <- plot_ly(tmp4, labels = ~`Have you observed Lesser yellow headed vultures in your farm? How often?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))
f4
```

#### Greater-yellow

```{r}
tmp5 <- farm_db %>% group_by( `Have you observed greater yellow headed vultures in your farm? How often?` ) %>% summarise(freq=n())
f5 <- plot_ly(tmp5, labels = ~`Have you observed greater yellow headed vultures in your farm? How often?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))
f5
```

### What animals have you seen the vultures feeding on?

Simplified version below:

```{r}
res <- list()
res[[1]] <- data.frame(animal="Cattle", n=sum(farm_db$Cow_carcass, na.rm=T))
res[[2]] <- data.frame(animal="Horse", n=sum(farm_db$Horse_carcass, na.rm=T))
res[[3]] <- data.frame(animal="Pig", n=sum(farm_db$Pig_carcass, na.rm=T))
res[[4]] <- data.frame(animal="Chicken", n=sum(farm_db$Chicken_carcass, na.rm=T))

res <- bind_rows(res)

fig <- plot_ly(
  x = res$animal,
  y = res$n,
  type = "bar"
)

fig


```

### Advantages of vultures on your farm?

```{r}
#farm_db$`What do you see the advantages of vultures on your farm?`
tmp <- farm_db %>% group_by(`What do you see the advantages of vultures on your farm?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What do you see the advantages of vultures on your farm?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```


### Disadvantages of vultures on your farm?

```{r}
#farm_db$`What do you see the disadvantages of vultures on your farm?`
tmp <- farm_db %>% group_by(`What do you see the disadvantages of vultures on your farm?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`What do you see the disadvantages of vultures on your farm?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Changes in vulture populations in your area in the past 5 years?

```{r}
#farm_db$`Have you noticed any changes in vulture populations in your area in the past 5 years.`
tmp <- farm_db %>% group_by(`Have you noticed any changes in vulture populations in your area in the past 5 years.`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Have you noticed any changes in vulture populations in your area in the past 5 years.`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```

### Interested in monitoring your farms vultures in the future?

```{r}
#farm_db$`Would you be interested in monitoring your farms vultures in the future?`
tmp <- farm_db %>% group_by(`Would you be interested in monitoring your farms vultures in the future?`  ) %>% summarise(freq=n())

fig <- plot_ly(tmp, labels = ~`Would you be interested in monitoring your farms vultures in the future?`   , values = ~freq, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hoverinfo = 'text',
        text = ~paste(freq))

fig
```
